% =============================================================================
% Executive Summary
% =============================================================================
\section*{Executive Summary}\label{sec:executive-summary}
\addcontentsline{toc}{section}{Executive Summary}

This paper presents the complete empirical results of the International
Sovereignty Index (ISI) applied to the 27~member states of the European
Union, using vintage-2024 data processed under methodology v1.0.  The
index measures the concentration of bilateral external dependencies
across six structural axes.  Higher scores indicate higher concentration
of counterparty exposure.  The analysis goes substantially beyond
descriptive ranking to provide a structurally decomposed,
variance-explained, rank-stable, and statistically interrogated
empirical system.  The principal findings are:

\begin{itemize}[leftmargin=2em]

\item \textbf{Compressed composite distribution.}
  The ISI composite ranges from \num{0.23567126} (France) to
  \num{0.51748148} (Malta), a span of \num{0.28181022} on the unit
  interval.  The population standard deviation is \num{0.06992188},
  yielding a coefficient of variation of 0.2030.  The interquartile
  range is \num{0.09881618} (P25~= \num{0.29179051}, P75~=
  \num{0.39060669}).  The Gini coefficient is \num{0.115463},
  indicating a relatively egalitarian distribution.
  See \cref{tab:summary-stats,fig:histogram,fig:lorenz}.

\item \textbf{Tier compression is mathematically inevitable.}
  Under the four-tier classification, 23 of 27~countries (85.2\%)
  fall in the \emph{moderately concentrated} tier (ISI $\in [0.25,
  0.50)$).  Only Malta exceeds 0.50; three countries (Latvia,
  Slovakia, France) fall below~0.25.  The between-tier $\eta^2 =
  0.483$ indicates that the official scheme captures less than half
  the total dispersion; the within-tier range for the dominant
  category spans \num{0.20727287}, nearly four times the entire
  mildly-concentrated band.
  See \cref{tab:ranking,tab:compression-diagnostics,fig:rank-scatter}.

\item \textbf{Defence and logistics dominate cross-country variance.}
  A covariance-based variance decomposition, proved analytically from
  the linear aggregation rule, attributes 35.10\% of composite
  variance to the defence axis and 34.26\% to the logistics axis.
  Together these two axes account for 69.37\% of total composite
  variance.  The own-variance component accounts for 64.28\% of total
  composite variance; cross-covariances account for 35.72\%.
  Financial concentration contributes only 1.03\%.
  See \cref{tab:variance-decomp,tab:variance-decomp-detail,fig:axis-dominance-bar}.

\item \textbf{Axis dominance is concentrated.}
  In 16 of 27~countries, the defence axis records the highest
  axis-level score; in the remaining 11, logistics is highest.
  No country has its maximum score on the financial, energy,
  technology, or critical-inputs axis.  The mean defence-to-second
  gap for defence-dominant countries is \num{0.183660}.
  See \cref{tab:axis-dominance,tab:defence-structural}.

\item \textbf{Defence axis exhibits oligopsonistic concentration.}
  Five EU member states record defence axis scores exceeding 0.80,
  and 59.3\% of countries have defence as their dominant axis.  The
  distribution is positively skewed, and Slovakia records the only
  zero score in the entire dataset.  This structural pattern reflects
  oligopsonistic procurement markets in which small
  buyers face few feasible suppliers.
  See \cref{tab:defence-structural,tab:defence-gap}.

\item \textbf{Three moderate cross-axis correlations; only two
  statistically significant.}
  Of 15~pairwise Pearson correlations, only two are significant at
  the 5\% level ($t_{0.025,25} = 2.060$): critical inputs--logistics
  ($r = 0.549$, $t = 3.280$) and energy--technology ($r = 0.530$,
  $t = 3.124$).  The remaining 13~pairs---including all correlations
  involving the financial axis---are not significant, establishing
  that the six axes capture largely distinct dimensions.
  See \cref{tab:corr-pearson,tab:corr-tstat,fig:corr-heatmap,fig:corr-network}.

\item \textbf{Rank ordering robust at extremes, fragile in mid-tier.}
  Under leave-one-axis-out (LOO) analysis, Spearman~$\rho$ ranges
  from 0.833 (defence excluded) to 0.993 (financial excluded); Kendall
  $\tau$ from 0.675 (defence) to 0.943 (financial, energy).  Under
  Dirichlet weight perturbation (10\,000 draws, $\alpha_j = 10$),
  Malta ($\sigma < 0.5$~ranks) and France ($\sigma < 0.5$~ranks) are
  structurally locked, while the Netherlands ($\sigma = 3.07$) is the
  most volatile.  Under $\pm10\%$ axis perturbation, 14 of 27
  countries are rank-invariant.
  See \cref{tab:loo-summary,tab:dirichlet-volatility,tab:rank-elasticity}.

\item \textbf{Financial concentration is orthogonal to goods-trade
  concentration.}
  The financial axis is nearly uncorrelated with every other axis
  ($|r| \leq 0.22$, all $|t| < 1.15$).  Its LOO removal produces
  $\rho = 0.993$, the highest stability of any axis.  Its variance
  contribution is 1.03\%.  Its bivariate $R^2$ with rank is 0.025, and
  its partial correlation with rank (controlling for all other axes) is
  0.421.  This establishes that financial-dependence patterns are
  structurally distinct from physical-trade dependence patterns.
  See \cref{tab:axis-rank-contrib,tab:corr-tstat}.

\item \textbf{Compensability masks profile heterogeneity.}
  The arithmetic-mean aggregation rule allows high scores on one axis
  to compensate for low scores on others.  The Netherlands exhibits the
  highest compensation gap (max axis $= 0.960$, composite $= 0.335$,
  gap $= 0.624$, CV~$= 0.880$), followed by Italy (gap $= 0.547$)
  and Bulgaria (gap $= 0.520$).
  See \cref{tab:compensability}.

\item \textbf{Geographic and structural correlates.}
  Island and near-island economies (Malta, Cyprus, Ireland) average
  ISI~$= 0.471$, significantly above the continental mean of 0.329.
  Large-population countries ($> 20$M) average 0.298; small-population
  countries ($< 5$M) average 0.372.  Core EU-6 members average 0.312;
  peripheral members average 0.368.
  See \cref{tab:geo-correlates}.

\item \textbf{Slovakia's zero defence score.}
  Slovakia records a defence axis score of \num{0.00000000}, the
  only zero score in the dataset.  This arises mechanically from
  the absence of bilateral supplier entries for Slovakia in the
  SIPRI arms-transfer data under methodology v1.0.  It does not
  imply the absence of defence procurement.
  See \cref{sec:country-profiles,tab:defence-gap}.

\item \textbf{Defence-dominance stress test.}
  Three counterfactual composites---without defence, without
  logistics, and without both---demonstrate that the ranking
  retains structural coherence even when the two highest-variance
  axes are excluded (without both: $\rho = 0.654$, $R^2 = 0.428$).
  The ISI is not reducible to a ``defence index.''
  See \cref{tab:defence-stress-test}.

\item \textbf{Formal structural typology.}
  Countries are classified by their dominance gap $\Delta_i$
  into single-axis dominated ($\Delta_i \geq 0.30$; 3~countries,
  all defence-dominant), dual-axis concentrated
  ($0.10 \leq \Delta_i < 0.30$; 8~countries), and balanced
  ($\Delta_i < 0.10$; 16~countries).  Profile polarisation
  predicts rank volatility under weight perturbation.
  See \cref{tab:structural-typology}.

\item \textbf{PCA component loadings.}
  The first principal component (31.7\% of variance) loads on the
  four goods-trade axes; the second (26.4\%) loads on defence and
  logistics; the third (18.1\%) isolates the financial axis.
  PCA-derived weighting would suppress the defence axis---the
  single most variance-contributing dimension---reinforcing the
  case for equal weighting.
  See \cref{tab:pca-loadings}.

\end{itemize}
