% =============================================================================
% Section 1: Introduction & Contribution
% =============================================================================
\section{Introduction and Contribution}\label{sec:introduction}

\subsection{Motivation}

Composite indicators that measure structural features of national
economies occupy a central position in institutional research.  The
OECD Handbook on Constructing Composite Indicators establishes the
standard methodological framework: variable selection, normalisation,
weighting, and aggregation, each step carrying design choices that
shape the information content of the final
score~\cite{oecd_handbook_2008}.  Saisana and
Tarantola~\cite{saisana2005} demonstrate that rigorous sensitivity
and uncertainty analysis is indispensable for any composite index
that informs policy.  The International Sovereignty Index (ISI)
adheres to both standards and extends them to a domain that has
received insufficient quantitative treatment: \emph{the concentration
of bilateral external dependencies across strategically distinct
domains}.

The intellectual foundations of the ISI rest on three pillars.
First, the industrial-organisation theory of market concentration,
where the Herfindahl--Hirschman Index (HHI) provides a closed-form
summary of distributional inequality among market
participants~\cite{herfindahl1950,hirschman1964}.  HHI is
well-understood theoretically---it equals the inverse of the
``numbers equivalent'' of equally sized suppliers---and its
behavioural interpretation in terms of mark-up and price--cost
margins is established in the antitrust literature.  The ISI applies
this logic at the country--partner level rather than the
firm--market level.

Second, the literature on supply-chain vulnerability, which expanded
substantially after the semiconductor shortages of 2020--2021 and
the geopolitical disruptions following the 2022 Russia--Ukraine
conflict~\cite{ec_critical_raw}.  The European Commission's Critical
Raw Materials assessments, the US Executive Order on Supply Chain
Resilience, and OECD analyses of trade concentration in critical
minerals all employ concentration metrics---typically HHI or
$CR_k$ ratios---to identify bottleneck dependencies.  The ISI
formalises and generalises these sector-specific exercises into a
multi-axis composite framework.

Third, the methodological debate on equal versus differential
weighting in composite indicators.  Equal weighting is the most
common default in practice (used by the Human Development Index,
the Global Innovation Index, and the OECD Better Life Index, among
others), but it is not neutral: it implicitly assumes that all
dimensions are equally important and that the analyst has no
empirical or normative basis for asymmetric treatment.  The ISI
adopts equal weighting as its baseline specification and subjects
this choice to explicit statistical interrogation through Dirichlet
weight perturbation (\cref{sec:dirichlet}) and variance
decomposition (\cref{sec:variance-decomp}).

The ISI does not measure sovereignty in a political or legal sense.
It quantifies a structural feature---counterparty
concentration---that is a prerequisite for informed policy analysis of
external dependence.  A high ISI score indicates that a country's
external flows in a given domain are heavily concentrated among few
partners; a low score indicates dispersion across many partners.  The
index makes no normative claim about whether concentration is
advantageous or disadvantageous.  Whether high concentration
translates into political vulnerability, economic fragility, or
coercive leverage is a distinct empirical question that lies beyond
the ISI's measurement scope.

\subsection{Scope of this paper}

This paper---the second in the ISI series---presents the full
empirical results of the first ISI application: 27~EU member states,
vintage-2024 data, methodology v1.0.  The methodological foundations
(axis definitions, normalisation procedures, data sources, aggregation
rule, integrity chain) are documented in Paper~1~\cite{isi_paper1}.
The present paper is self-contained for readers interested in the
empirical findings; cross-references to Paper~1 are provided where
construction details are relevant to interpretation.

Relative to a conventional ``results'' paper in the composite-indicator
literature, this paper provides substantially deeper analytical content.
Beyond reporting the ranking and basic distributional statistics, the
paper undertakes a formal variance decomposition with proof, systematic
robustness exercises with both Spearman and Kendall rank-order
statistics, rank-elasticity analysis, axis contribution decomposition,
classification compression diagnostics, structural analysis of the
defence axis, a defence-dominance stress test, PCA-based dimensionality
diagnostics with component loadings, compensability analysis, formal
structural typology, and geographic correlate tabulation.  Each
analytical component interrogates the ISI composite from a different
angle and provides the evidence required for the reader to form an
independent judgment about the index's informational content.

The specific contributions of this paper are:

\begin{enumerate}[leftmargin=2em]
  \item A complete composite and axis-level ranking of all 27~EU
    member states with eight-decimal precision, accompanied by
    classification labels and distributional diagnostics (histogram,
    ECDF, Lorenz curve, Gini coefficient).
  \item A formal variance decomposition---proved analytically from the
    linear aggregation rule---decomposing the composite variance into
    own-variance and cross-covariance components, with percentage
    contributions for each axis.
  \item A leave-one-axis-out (LOO) sensitivity analysis with both
    Spearman~$\rho$ and Kendall~$\tau$, accompanied by a full $27
    \times 6$ rank-displacement matrix.
  \item A rank-elasticity analysis quantifying the sensitivity of each
    country's rank to marginal perturbations in individual axis scores.
  \item An axis-contribution analysis providing bivariate $R^2$ and
    partial correlations between each axis and the composite rank.
  \item Classification compression diagnostics comparing the
    official four-tier scheme with data-driven tercile and quartile
    alternatives, reporting within-tier variance and $\eta^2$.
  \item A structural analysis of the defence axis examining its
    distributional properties, oligopsonistic character, and
    country-level defence-to-max-axis gaps.
  \item A defence-dominance stress test constructing counterfactual
    composites without the defence axis, without the logistics axis,
    and without both, to demonstrate that the ranking retains
    structural coherence independently of the two highest-variance
    axes.
  \item A formal structural typology classifying countries by
    dominance gap ($\Delta_i$) into single-axis dominated, dual-axis
    concentrated, and balanced profiles.
  \item A cross-axis correlation analysis with formal $t$-statistics,
    Bonferroni-corrected multiple-testing assessment,
    eigenvalue decomposition, PCA component loadings with
    interpretation, and network representation.
  \item A compensability analysis with formal gap, range, and CV
    metrics documenting the degree to which the arithmetic-mean
    aggregation rule permits cross-axis compensation.
  \item An assessment of geographic and structural correlates (island
    status, population size, core/periphery classification).
  \item A four-part robustness programme: LOO, z-score
    standardisation, Dirichlet weight perturbation (Monte Carlo), and
    winsorisation.
  \item Radar-chart profiles for eight selected countries illustrating
    the heterogeneity of axis-level patterns.
  \item A reproducibility protocol and computational appendix enabling
    full replication from the provided snapshot.
\end{enumerate}

\subsection{Interpretation conventions}

Throughout this paper, the following conventions apply:

\begin{itemize}[leftmargin=2em]
  \item Higher ISI composite $\Rightarrow$ higher concentration
    $\Rightarrow$ greater structural dependence on a small set of
    counterparties.
  \item Rank~1 = most concentrated country; rank~27 = least
    concentrated.
  \item The terms ``most concentrated'' and ``least concentrated'' are
    descriptive.  They are not synonymous with ``worst'' and ``best.''
  \item Correlation between axes does not imply causation.  Where
    plausible structural interpretations exist, they are stated as
    hypotheses, not conclusions.
  \item Population standard deviation ($N$ denominator) is used
    throughout because the 27~EU member states constitute the full
    target population, not a sample drawn from a larger universe.
  \item Statistical significance of correlation coefficients is
    assessed via the $t$-test: $t = r\sqrt{(N-2)/(1-r^2)}$, with
    $N = 27$, df~$= 25$, and a two-sided critical value of
    $t_{0.025,25} = 2.060$ at the 5\% level.
\end{itemize}

\subsection{Structure}

The remainder of this paper is organised as follows.
\Cref{sec:data-construction} summarises the data sources, aggregation
rule, and classification thresholds.
\Cref{sec:composite-results} presents the composite distribution,
full ranking, tier-compression analysis, and classification
compression diagnostics.
\Cref{sec:axis-diagnostics} provides axis-level diagnostics including
dispersion, dominance, formal structural typology, variance
decomposition with formal proof, rank contribution analysis, defence
axis structural analysis, and a defence-dominance stress test.
\Cref{sec:cross-axis} examines the pairwise correlation structure
with $t$-statistics, Bonferroni-corrected multiple testing, PCA
component loadings, and eigenvalue decomposition.
\Cref{sec:robustness} reports the robustness programme: LOO with
Kendall $\tau$, rank-displacement matrices, z-score standardisation,
Dirichlet weight perturbation, winsorisation, and rank-elasticity
analysis.
\Cref{sec:compensability} presents the compensability analysis.
\Cref{sec:country-profiles} presents radar-chart profiles for eight
selected countries and geographic correlates.
\Cref{sec:discussion} discusses interpretation boundaries and
limitations.
\Cref{sec:conclusion} concludes.
Appendices~A--D provide full tables, robustness details, the
reproducibility protocol, and the computational code listing.
