% =============================================================================
% Section 6: Robustness Programme
% =============================================================================
\section{Robustness Programme}\label{sec:robustness}

This section subjects the ISI composite ranking to six distinct
perturbation exercises.  Each test isolates a different source of
methodological sensitivity: (i)~axis inclusion, (ii)~standardisation
choice, (iii)~weight specification, (iv)~outlier treatment,
(v)~rank displacement under LOO, and (vi)~rank elasticity to
axis-level perturbation.  The baseline is the ISI~v1.0 ranking
(\cref{tab:ranking}).

% -------------------------------------------------------------------------
\subsection{Leave-one-axis-out (LOO)}\label{sec:loo}
% -------------------------------------------------------------------------

For each of the six axes in turn, we recompute the composite as the
simple average of the remaining five axes and re-rank the~27 countries.
We report both the Spearman rank correlation~$\rho$ and the
Kendall~$\tau$ concordance coefficient between the reduced ranking
and the baseline ranking.

\input{tables/T09_loo_summary}

\Cref{tab:loo-summary} reports the results.  Key findings:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Financial axis removal} ($\rho = 0.993$,
    $\tau = 0.943$).
    Dropping the financial axis barely perturbs the ranking.
    This result follows directly from its low variance contribution
    (1.03\%, \cref{tab:variance-decomp}) and near-zero
    correlations with other axes.

  \item \textbf{Energy axis removal} ($\rho = 0.992$,
    $\tau = 0.943$).
    Similarly negligible rank disruption despite the moderate
    correlation of energy with technology ($r = 0.53$,
    \cref{tab:corr-pearson}).

  \item \textbf{Technology axis removal} ($\rho = 0.968$,
    $\tau = 0.886$).
    Modest rank perturbation.  Ireland, the highest-scoring
    country on the technology axis (0.587), loses relative
    position when this axis is removed.

  \item \textbf{Defence axis removal} ($\rho = 0.833$,
    $\tau = 0.675$).
    The most disruptive single omission.  Defence contributes
    35.10\% of composite variance and has the highest absolute
    dispersion ($\sigma = 0.226$).  Its removal reshuffles the
    middle ranks substantially, particularly for countries like
    the Netherlands (defence score = 0.960) and Bulgaria
    (0.861) whose composite positions depend heavily on
    their defence concentration.  The Kendall~$\tau$ of 0.675
    indicates that roughly one-third of all pairwise rank
    comparisons reverse when defence is omitted.

  \item \textbf{Critical inputs removal} ($\rho = 0.941$,
    $\tau = 0.846$).
    Moderate disruption, consistent with its 20.92\% variance
    contribution.

  \item \textbf{Logistics axis removal} ($\rho = 0.926$,
    $\tau = 0.789$).
    Moderate disruption, consistent with its 34.26\% variance
    contribution.  Malta and Cyprus, the top-two countries, are
    most affected because both score 1.00 and 0.94 respectively
    on logistics.
\end{itemize}

The concordance between Spearman~$\rho$ and Kendall~$\tau$ is
high across all six exercises: the ordinal relationship
$\text{Defence} < \text{Logistics} < \text{Critical} < \text{Tech}
< \text{Energy} \approx \text{Financial}$ holds for both statistics.
The LOO results are therefore not artefacts of a particular
rank-correlation metric.

The full country-by-country LOO ranking table is provided in
\cref{tab:loo-detail} (\cref{app:robustness}).

% -------------------------------------------------------------------------
\subsection{Rank displacement matrix}\label{sec:rank-displacement}
% -------------------------------------------------------------------------

To identify \emph{which} countries are most affected by each axis
omission, \cref{tab:rank-displacement} reports the signed rank change
$\Delta r_i^{(-j)} = r_i^{(-j)} - r_i^{\text{baseline}}$ for each
country~$i$ and omitted axis~$j$.  A positive $\Delta r$ indicates
worsening (higher rank number = further from rank~1).

\input{tables/TE2_rank_displacement}

The following displacement patterns are analytically significant:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Maximum displacements.}  The largest absolute rank
    changes when defence is omitted are observed for Italy
    ($|\Delta r| = 10$), Slovakia ($|\Delta r| = 10$), and the
    Netherlands ($|\Delta r| = 10$).  These countries have highly
    concentrated defence-import profiles, and their composite
    position depends substantially on this single axis.

  \item \textbf{Bulgaria and Finland.}  Bulgaria shows displacement
    of~$|\Delta r| = 9$ on defence removal; Finland of~$|\Delta r| = 8$.
    Both sit in the compressed mid-range (\cref{sec:compression-diagnostics})
    where small score changes translate into large rank shifts.

  \item \textbf{Asymmetric vulnerability.}  Not all countries with
    high composite scores are sensitive to the same axis.  For instance,
    Malta (rank~1) is sensitive primarily to logistics removal
    ($\Delta r \geq 2$), while Cyprus (rank~2) is sensitive to both
    logistics and critical inputs, reflecting their distinct strategic
    profiles.

  \item \textbf{Stable anchors.}  France (rank~27) shows
    $|\Delta r| \leq 1$ under \emph{all} axis omissions, establishing
    its position as the most diversified and hence least sensitive
    country.
\end{itemize}

% -------------------------------------------------------------------------
\subsection{Z-score standardisation}\label{sec:zscore}
% -------------------------------------------------------------------------

The baseline ISI uses raw (min--max bounded) axis scores.  An
alternative is to standardise each axis to zero mean and unit variance
before aggregation.  Z-score standardisation implicitly upweights
axes with higher absolute dispersion (because a one-unit change in
a high-$\sigma$ axis maps to a smaller z-increment than a one-unit
change in a low-$\sigma$ axis after rescaling).

We compute:
\begin{equation}\label{eq:zscore}
  z_{ij} = \frac{A_{ij} - \bar{A}_j}{\sigma_j}\,,
  \qquad
  \text{ISI}^{z}_i = \frac{1}{6}\sum_{j=1}^{6} z_{ij}\,,
\end{equation}
and re-rank the countries.

\input{tables/T10_zscore_ranks}

\Cref{tab:zscore-ranks} compares the z-score-based ranking with the
baseline.  The Spearman rank correlation is $\rho = 0.896$, the
Kendall~$\tau = 0.726$---lower than any single LOO perturbation except
defence removal.  This indicates that the implicit dispersion-weighting
embedded in z-score standardisation produces a meaningfully different
ranking.

The largest rank changes under z-score standardisation include:
\begin{itemize}[leftmargin=2em]
  \item Countries with extreme scores on high-CV axes (technology,
    critical inputs) gain or lose relative position depending on
    whether the z-score upweighting favours their profile.
  \item Countries whose composite depends predominantly on the
    defence axis (which has moderate CV but the largest absolute
    $\sigma$) are less affected, because z-score standardisation
    compresses defence scores toward the mean.
\end{itemize}

The result underscores that the choice of standardisation is not
innocuous.  The ISI~v1.0 deliberately avoids z-score normalisation
because the underlying data (HHI-normalised scores) already inhabit
a common $[0, 1]$ scale, making further rescaling conceptually
unnecessary and potentially distortive.

% -------------------------------------------------------------------------
\subsection{Dirichlet weight perturbation (Monte Carlo)}\label{sec:dirichlet}
% -------------------------------------------------------------------------

The baseline ISI assigns equal weight ($w_j = 1/6$) to each axis.
To assess sensitivity to this choice, we draw $K = 10{,}000$
weight vectors from a symmetric $\text{Dirichlet}(\alpha = 1)$
distribution (uniform over the 5-simplex), compute the weighted
composite for each draw, rank the countries, and record the rank
distribution.

\input{tables/T11_dirichlet_volatility}

\Cref{tab:dirichlet-volatility} reports the mean and standard
deviation of each country's rank across the 10\,000 draws, along
with the 5th and 95th percentile ranks.  Aggregate summary
statistics:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Mean Spearman~$\rho$} between each draw's ranking
    and the baseline: \num{0.979} (std~=~\num{0.017}).
  \item \textbf{Minimum~$\rho$}: \num{0.863}; \textbf{Maximum~$\rho$}:
    \num{1.000}.
  \item In more than 95\% of draws, $\rho > 0.95$, indicating that
    the ranking is robust to substantial weight variation.
\end{itemize}

\textbf{Most rank-volatile countries:}
\begin{enumerate}[leftmargin=2em]
  \item Netherlands ($\sigma_{\text{rank}} = 3.07$) --- rank 15 on
    average, but ranges from 3 to 25 across draws.  This extreme
    volatility reflects the Netherlands' highly polarised profile:
    defence = 0.960 but financial = 0.121 and technology = 0.125.
  \item Finland ($\sigma_{\text{rank}} = 2.39$) --- sensitive to
    the weight placed on critical inputs (highest score: 0.547) vs
    defence (low score: 0.390).
  \item Estonia ($\sigma_{\text{rank}} = 2.21$).
  \item Italy ($\sigma_{\text{rank}} = 2.17$).
  \item Bulgaria ($\sigma_{\text{rank}} = 2.03$).
\end{enumerate}

\textbf{Most rank-stable countries:}  Malta ($\sigma_{\text{rank}} <
0.5$) and France ($\sigma_{\text{rank}} < 0.5$) are the most stable,
reflecting their positions at the extreme top and bottom of the
ranking where re-weighting has limited scope to shift them.

% -------------------------------------------------------------------------
\subsection{Winsorization at the 95th percentile}\label{sec:winsorization}
% -------------------------------------------------------------------------

Two axes---defence and logistics---contain country scores at the
boundary value of~1.0.  To assess whether these extreme values
drive the ranking, we cap both axes at their respective 95th
percentiles:

\begin{itemize}[leftmargin=2em]
  \item Defence: P95 cap = \num{0.94786072}.
  \item Logistics: P95 cap = \num{0.89794180}.
\end{itemize}

Any country-axis score exceeding the cap is replaced by the cap value.
The composite is then recomputed and re-ranked.

\input{tables/T12_winsorization}

\Cref{tab:winsorization} compares the winsorized ranking with the
baseline.  The Spearman rank correlation is $\rho = 1.000$: the
ranking is entirely preserved.  This means that no country's ordinal
position depends on having an axis score above the 95th percentile.
The result provides strong evidence that the ranking is not an
artefact of boundary effects in the defence or logistics axes.

% -------------------------------------------------------------------------
\subsection{Rank elasticity to axis perturbation}\label{sec:rank-elasticity}
% -------------------------------------------------------------------------

A different form of sensitivity analysis asks: for each country,
does a $\pm 10\%$ perturbation of \emph{any single axis score}
change the country's composite rank?

Formally, for country~$i$ and axis~$j$, we compute the perturbed
composite
\begin{equation}\label{eq:rank-elasticity}
  \text{ISI}_i^{(j,\pm)} = \frac{1}{6}\left[
    \sum_{k \neq j} A_{ik} + A_{ij}(1 \pm 0.10)
  \right],
\end{equation}
re-rank all 27~countries, and record whether country~$i$'s rank
changes.  The total number of rank-changing perturbations (out of
12~possible: 6~axes $\times$ 2~directions) measures the country's
\emph{rank elasticity}.

\input{tables/T13_rank_elasticity}

\Cref{tab:rank-elasticity} reports the results.  The 27~countries
split into two groups:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Rank-elastic countries} (13 of 27):
    at least one axis perturbation changes the composite rank.
    Croatia, Finland, and Sweden each show total rank changes of~2,
    meaning that perturbation on a single axis is sufficient to shift
    their ranking by one or more positions.

  \item \textbf{Rank-locked countries} (14 of 27):
    no axis perturbation of $\pm 10\%$ changes the composite rank.
    Malta (rank~1), France (rank~27), and several mid-range countries
    with score gaps large enough to absorb the perturbation are
    entirely insensitive.
\end{itemize}

The rank-elastic countries are concentrated in the compressed
mid-range of the composite distribution
(\cref{sec:compression-diagnostics}), where inter-country score
differences are smallest.  The mid-range ranking
is intrinsically fragile: even modest data revisions or methodological
adjustments can reshuffle countries in the 0.30--0.40 ISI range.

Conversely, rank-locked countries are predominantly those at the
distributional extremes---the highest and lowest ISI scores---or
countries whose score gap to their nearest neighbour exceeds the
perturbation magnitude.  The countries that are most policy-relevant
(the most and least concentrated) are therefore also the most
robustly ranked.

% -------------------------------------------------------------------------
\subsection{Robustness summary}\label{sec:robustness-summary}
% -------------------------------------------------------------------------

\Cref{tab:robustness-summary} consolidates the six robustness
exercises.

\begin{table}[htbp]
\centering
\caption{Robustness programme summary.  Spearman~$\rho$ and
  Kendall~$\tau$ between each perturbed ranking and the ISI~v1.0
  baseline.}
\label{tab:robustness-summary}
\begin{tabular}{@{}llS[table-format=1.3]S[table-format=1.3]l@{}}
\toprule
\textbf{Exercise} & \textbf{Perturbation}
  & {$\boldsymbol{\rho}$} & {$\boldsymbol{\tau}$}
  & \textbf{Verdict} \\
\midrule
LOO (Financial)   & Drop axis 1 & 0.993 & 0.943 & Stable \\
LOO (Energy)      & Drop axis 2 & 0.992 & 0.943 & Stable \\
LOO (Technology)  & Drop axis 3 & 0.968 & 0.886 & Stable \\
LOO (Defence)     & Drop axis 4 & 0.833 & 0.675 & Moderate sensitivity \\
LOO (Critical)    & Drop axis 5 & 0.941 & 0.846 & Stable \\
LOO (Logistics)   & Drop axis 6 & 0.926 & 0.789 & Stable \\
Z-score           & Re-standardise & 0.896 & 0.726 & Moderate sensitivity \\
Dirichlet MC      & Random weights & 0.979 & {--}  & Stable (mean) \\
Winsorization     & Cap at P95 & 1.000 & 1.000 & Fully stable \\
\midrule
\multicolumn{2}{@{}l}{Rank elasticity ($\pm 10\%$)}
  & \multicolumn{3}{@{}l}{13/27 sensitive; 14/27 locked} \\
\bottomrule
\end{tabular}
\end{table}

The ranking is robust to outlier capping (winsorization) and to
random weight perturbation (Dirichlet MC).  The two sources of
moderate sensitivity are (i)~defence axis removal ($\rho = 0.833$,
$\tau = 0.675$) and (ii)~z-score standardisation ($\rho = 0.896$,
$\tau = 0.726$).  Both are structural: defence is the single most
dispersed and variance-contributing axis, and z-score standardisation
introduces implicit dispersion-based reweighting.  Neither constitutes
a vulnerability in the index design; rather, they highlight the
importance of the defence axis to cross-country differentiation and
the deliberate choice to avoid dispersion-sensitive normalisation.

The rank elasticity exercise further refines the picture: while 13
countries are sensitive to $\pm 10\%$ axis-level perturbations, these
are concentrated in the compressed mid-range.  The extreme-ranked
countries (Malta, Cyprus, France, Germany) are entirely stable.  This
convergence of evidence---across LOO, Monte Carlo, winsorization,
and elasticity tests---supports the conclusion that the ISI~v1.0
ranking is robust at the distributional extremes and fragile only in
the compressed middle tier where inter-country score differences are
small and inherently difficult to resolve with any aggregation method.
