% =============================================================================
% Section 8: Discussion
% =============================================================================
\section{Discussion}\label{sec:discussion}

The empirical results converge on five structural conclusions about
the ISI's informational architecture, each with distinct implications
for measurement practice and policy application.  This section
presents those conclusions, derives their measurement and policy
implications, and articulates the methodological limitations that
bound their interpretation.

\subsection{Five structural conclusions}

\paragraph{Conclusion~1: The ISI is dominated by defence and logistics
variance.}
The variance decomposition (\cref{sec:variance-decomp}) establishes that
the defence axis contributes 35.10\% and the logistics axis 34.26\%
of total composite variance, jointly accounting for 69.37\%.  The
own-variance component constitutes 64.28\% and the cross-covariance
component 35.72\% (\cref{tab:variance-decomp-detail}).  Cross-country
ranking differences are overwhelmingly determined by two of the six
axes; the remaining four axes (financial, energy, technology, critical
inputs) together contribute less than one-third of the composite's
discriminatory power.

This dominance is not an artefact of scaling or weighting.  Because
all axes inhabit a common $[0,1]$ scale and receive equal weight,
the variance shares reflect genuine differences in cross-country
dispersion: defence has $\sigma = 0.226$ and logistics $\sigma = 0.186$,
while the financial axis has $\sigma = 0.037$.  The multiple
regression analysis (\cref{sec:axis-rank}) confirms this: defence and
logistics have the highest bivariate $R^2$ values (0.394 and 0.491,
respectively) and partial correlation coefficients (0.940 and 0.809)
with the composite, and a six-axis regression achieves $R^2 = 0.968$.

The PCA loadings (\cref{tab:pca-loadings}) provide independent
structural confirmation: defence does not load on the broadest
goods-trade factor (PC1) but instead defines its own latent
dimension (PC2) together with logistics.  This means that the
defence--logistics variance dominance operates through a
structurally distinct channel---oligopsonistic procurement
concentration---that is independent of the goods-trade
diversification patterns captured by the other four axes.

The defence-dominance stress test (\cref{sec:defence-stress-test})
directly addresses the implied objection.  Removing both defence and
logistics reduces rank concordance to $\rho = 0.654$ but preserves
42.8\% of baseline rank variance, demonstrating that the remaining
four axes carry independent discriminatory content.  The ISI is not
reducible to a ``defence index''; it is an index in which defence
concentration is the single most important axis of variation, a
finding that reflects the empirical structure of European strategic
dependencies.

\paragraph{Conclusion~2: Rank ordering is robust at extremes, fragile
in the compressed mid-tier.}
The robustness programme (\cref{sec:robustness}) demonstrates that
the ISI ranking is stable under winsorization ($\rho = 1.000$),
Dirichlet weight perturbation (mean $\rho = 0.979$), and five of
six LOO exercises ($\rho \geq 0.926$).  Only defence-axis removal
($\rho = 0.833$, $\tau = 0.675$) and z-score standardisation
($\rho = 0.896$, $\tau = 0.726$) produce moderate rank disruption.

The rank elasticity analysis (\cref{sec:rank-elasticity}) establishes that
13 of 27 countries are sensitive to $\pm 10\%$ axis perturbations, but
these are concentrated in the composite range 0.30--0.40 where
inter-country score gaps are smallest.  The extreme-ranked countries
(Malta, Cyprus, France, Germany) are entirely stable.  This pattern
is consistent with the compression diagnostic ($\eta^2 = 0.483$,
\cref{sec:compression-diagnostics}): the four-tier classification
explains only 48.3\% of within-composite variance, confirming that
the mid-range is genuinely compressed and inherently difficult to
resolve with \emph{any} aggregation method.

The structural typology (\cref{sec:structural-typology}) sharpens
this finding: the sixteen balanced-profile countries ($\Delta_i < 0.10$)
exhibit lower rank volatility under Dirichlet perturbation than the
three single-axis-dominated countries ($\Delta_i \geq 0.30$).
Profile polarisation is both a source of mid-tier rank fragility and
a predictor of which countries' positions are most sensitive to
methodological choices.

\paragraph{Conclusion~3: Classification compression is a distributional
consequence, not a design flaw.}
The compression diagnostics (\cref{tab:compression-diagnostics}) show
that the four-tier ISI classification (unconcentrated, mildly,
moderately, highly concentrated) achieves $\eta^2 = 0.483$ using
thresholds at 0.15, 0.25, and 0.50.  Alternative classification
schemes---tercile-based
($\eta^2 = 0.890$) and quartile-based ($\eta^2 = 0.933$)---achieve
higher between-group separation but sacrifice the structurally
motivated HHI-analogue thresholds that provide interpretive anchoring
to the industrial-organisation literature.

The low $\eta^2$ under the ISI thresholds is a mathematical
consequence of the underlying data distribution.  Twenty-three of 27
countries fall in the ``moderately concentrated'' tier (0.25--0.50),
creating a crowded middle category with within-tier dispersion of
$\sigma = 0.053$.  Any fixed-threshold classification of a unimodal
distribution centred in the mid-range will produce this outcome.
The compression is an empirical regularity---EU-27 countries exhibit
similar aggregate concentration levels despite substantial axis-level
heterogeneity---not a failure of the classification scheme.

\paragraph{Conclusion~4: European defence markets exhibit
oligopsonistic concentration.}
The defence structural analysis (\cref{sec:defence-structural})
establishes that 16 of 27 EU member states (59.3\%) have the defence
axis as their single highest-scoring axis.  Five countries exceed
0.80, and one country (Malta) reaches the theoretical maximum of
1.000 (\cref{tab:defence-structural}).  The mean gap between the
defence score and the next-highest axis is 0.184
(\cref{tab:defence-gap}).  The structural typology
(\cref{sec:structural-typology}) identifies 3 countries as
single-axis dominated ($\Delta_i \geq 0.30$); in all three cases
the dominant axis is defence.

This concentration spans small (Malta, Cyprus), medium (Netherlands,
Bulgaria, Greece), and large (Italy) member states.  The pattern reflects
the oligopsonistic structure of European defence
procurement, in which a small number of bilateral supplier
relationships (predominantly US--EU) create high HHI values that are
structurally resistant to diversification through conventional trade
policy.  The defence-dominance stress test
(\cref{sec:defence-stress-test}) demonstrates that removing the
defence axis reduces rank concordance to $\rho = 0.833$---the
largest single-axis disruption in the robustness programme---while
removing both defence and logistics yields $\rho = 0.654$.  Defence
is the axis with the greatest structural influence on the composite.

\paragraph{Conclusion~5: Financial concentration is orthogonal to
goods-trade concentration.}
The correlation analysis (\cref{sec:cross-axis}) demonstrates that
the financial axis has near-zero Pearson correlations with all five
other axes ($|r| \leq 0.22$), with all $t$-statistics below 1.15
(\cref{tab:corr-tstat}).  The geographic correlates
(\cref{sec:geo-correlates}) confirm this: island status, population
size, and founding-member status---all strong predictors of
goods-trade concentration---have essentially no predictive power
for financial concentration.

This orthogonality has a structural interpretation.  The financial
axis measures the concentration of financial-service imports
(EBOPS classification), which depends on the structure of a country's
banking and insurance sectors rather than on its physical trade
geography.  A country can have highly concentrated goods imports
(high defence, logistics, energy scores) while maintaining
diversified financial-service imports, and vice versa.

The compensability analysis (\cref{sec:compensability}) further
illuminates this point: countries with high CV values (peaked
profiles) are overwhelmingly peaked on defence or logistics, not on
the financial axis.  The financial axis contributes only 1.03\% of
composite variance precisely because its cross-country dispersion
is so low ($\sigma = 0.037$).

\subsection{Measurement implications}

\paragraph{Equal weighting is defensible and empirically supported.}
The Dirichlet Monte Carlo exercise demonstrates that equal weighting
produces rankings within the 95\% confidence envelope of randomly
weighted composites (mean $\rho = 0.979$).  The eigenvalue analysis
(\cref{tab:eigenvalues}) reinforces this: no single principal
component explains more than 31.7\% of variance, so there is no
empirical basis for assigning dominant weight to any particular axis.
The PCA loadings (\cref{tab:pca-loadings}) further demonstrate that
data-driven (PC1-proportional) weighting would suppress the defence
axis---the single most variance-contributing dimension---because
defence loads on PC2 rather than PC1.  Equal weighting is therefore
not merely a convenient default but the specification most consistent
with the empirical covariance structure.

This finding aligns with the broader composite-indicator literature.
The OECD Handbook~\cite{oecd_handbook_2008} identifies equal weighting
as appropriate when axes are conceptually non-hierarchical and the
analyst has no empirical or normative basis for asymmetric treatment.
Both conditions hold for the ISI: the six axes represent conceptually
distinct domains of strategic dependence, and no external criterion
(e.g., GDP impact, vulnerability severity) has been validated to
justify differential weighting.  The equal-weight specification is
therefore a deliberate and defensible methodological choice, not an
arbitrary default.

\paragraph{Composite interpretation requires axis-level context.}
The compensability analysis (\cref{sec:compensability}) demonstrates
that the arithmetic-mean composite systematically compresses peak
vulnerabilities.  The Netherlands
(ISI = 0.335, defence = 0.960) and Estonia (ISI = 0.343, defence
= 0.356) have similar composites but structurally different risk
profiles---a distinction captured by the max--composite gap ($G$)
and the coefficient of variation (CV) but invisible in the scalar
composite.  Policy recommendations derived solely from the composite
score are therefore insufficient; the axis-level profile and the
structural typology (\cref{sec:structural-typology}) are necessary
complements.

\paragraph{Threshold-based classification is communication shorthand.}
The compression diagnostics establish that any threshold-based
classification of the ISI distribution will produce a crowded middle
tier ($\eta^2 = 0.483$ for the ISI scheme vs.\ $\eta^2 = 0.890$
for data-driven terciles).  The four-tier labels are retained as
communication shorthand for non-specialist audiences but should not
be treated as analytically binding categories.  Users of the ISI
should consult the continuous composite and axis-level scores for
work requiring fine-grained country comparisons.

\subsection{Policy implications}

The five structural conclusions above carry direct implications for
policy design:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Prioritise defence and logistics diversification.}
    The dominance of defence and logistics variance implies that
    EU strategic-autonomy policy should focus diversification
    efforts on these two domains, which together account for
    69.4\% of cross-country composite variation.
  \item \textbf{Present rankings with uncertainty bands.}
    Mid-tier ranking fragility implies that country positions
    in the 0.30--0.40 ISI range should be reported with
    confidence intervals or rank ranges derived from the
    Dirichlet perturbation exercise, rather than as point
    estimates.
  \item \textbf{Deploy axis-level dashboards.}
    The compensability findings argue for policy dashboards
    that display axis-level profiles alongside the composite,
    to avoid masking single-axis vulnerabilities behind an
    innocuous composite average.
  \item \textbf{Treat financial diversification independently.}
    The orthogonality of the financial axis indicates that
    financial-service diversification and goods-trade
    diversification can be pursued as independent policy
    tracks with distinct institutional mechanisms.
\end{itemize}

\subsection{Methodological limitations}\label{sec:limitations}

Several limitations constrain the interpretation of these results.
They are stated explicitly to delimit the index's evidential scope
and to guide the design of future extensions.

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Concentration as a proxy for vulnerability.}
    The ISI measures \emph{counterparty concentration}, which is
    a necessary but not sufficient condition for strategic
    vulnerability.  A country may have a highly concentrated
    supplier base composed entirely of stable, allied partners,
    in which case high concentration does not translate into
    coercive leverage or supply-disruption risk.  The ISI
    intentionally excludes partner-identity assessment in v1.0;
    incorporating geopolitical risk weighting is a planned
    extension.

  \item \textbf{Static single-vintage design.}
    The ISI~v1.0 is computed for a single reference year
    (vintage~2024).  Temporal dynamics---how concentration changes
    in response to policy interventions, geopolitical shocks, or
    trade-diversion effects---cannot be assessed without panel
    data.  A country's current rank may reflect transient
    conditions rather than persistent structural features.

  \item \textbf{Small-state structural amplification.}
    Small economies (Malta, Cyprus, Luxembourg) face inherently
    fewer bilateral trade partners, mechanically elevating their
    HHI scores.  The ISI captures this size effect faithfully---it
    \emph{is} a structural feature of small-state dependence---but
    users should be aware that high ISI scores for small states
    partly reflect market-size constraints rather than policy
    choices.

  \item \textbf{Re-export and transit-trade masking.}
    Trade statistics record the \emph{declared} partner country,
    which may be a re-export hub (e.g., the Netherlands, Belgium)
    rather than the ultimate origin.  This can understate the true
    concentration of ultimate-source dependence for countries whose
    imports transit through intermediary economies.  The magnitude
    of this effect varies by axis and is not corrected in v1.0.

  \item \textbf{Equal weighting as a normative choice.}
    Although the Dirichlet exercise demonstrates rank stability
    under random weighting, specific policy contexts exist in
    which defence concentration should receive higher weight than
    financial concentration.  The ISI framework accommodates
    differential weighting but does not prescribe it.  Users
    with domain-specific priors should apply the axis-level
    scores directly.

  \item \textbf{Data-window heterogeneity.}
    The defence axis uses a six-year rolling window (2019--2024)
    while all other axes use a three-year window (2022--2024).
    This reflects the lower frequency of arms-transfer
    flows but introduces temporal asymmetry: the defence axis
    may reflect structural patterns that pre-date the observation
    windows of other axes.

  \item \textbf{HHI as the sole concentration measure.}
    The axes use Herfindahl--Hirschman indices or comparable
    concentration statistics.  Alternative measures (entropy,
    Gini of trade shares, bilateral dependency ratios) might
    produce different rankings, particularly for countries near
    the median where score gaps are smallest.

  \item \textbf{Small $N$.}  With $N = 27$, many pairwise
    correlations fail to reach statistical significance even when
    point estimates are moderate (\cref{tab:corr-tstat}).  The
    geographic correlates (\cref{sec:geo-correlates}) are
    descriptive, not inferential, because the small sample
    precludes multivariate regression with more than a few
    covariates.
\end{enumerate}

\noindent These limitations are inherent to the v1.0 design and do
not invalidate the empirical findings.  They define the boundary
conditions within which the ISI results should be interpreted and
provide a structured agenda for methodological extension.
