% =============================================================================
% Section 9: Conclusion
% =============================================================================
\section{Conclusion}\label{sec:conclusion}

This paper has presented a comprehensive empirical analysis of the
International Sovereignty Index (ISI) v1.0 across the 27 member states
of the European Union.  The ISI aggregates six axis-level concentration
scores---financial services, energy, technology, defence, critical
inputs, and logistics---into a single composite indicator of strategic
import dependence.  The analysis has gone substantially beyond
descriptive ranking to provide a rigorous statistical characterisation
of the index's structure, robustness, and interpretive boundaries.

\subsection{Summary of contributions}

The paper makes the following analytical contributions:

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Complete distributional characterisation.}
    The ISI composite distribution is unimodal, moderately
    right-skewed (range 0.236--0.517, mean 0.344, $\sigma = 0.070$),
    with 23 of 27 countries in the ``moderately concentrated'' tier
    (\cref{sec:composite-results}).

  \item \textbf{Full variance decomposition with proof.}
    Defence (35.10\%) and logistics (34.26\%) jointly account for
    69.37\% of composite variance; own-variance contributes 64.28\%
    and cross-covariance 35.72\% (\cref{sec:variance-decomp}).

  \item \textbf{Leave-one-axis-out with Kendall concordance.}
    LOO analysis with both Spearman~$\rho$ and Kendall~$\tau$
    establishes that defence removal ($\rho = 0.833$, $\tau = 0.675$)
    is the most disruptive perturbation (\cref{sec:loo}).

  \item \textbf{Rank displacement matrix.}
    Country-by-axis signed rank changes identify Italy, Slovakia, and
    the Netherlands as the most displacement-sensitive countries
    (\cref{sec:rank-displacement}).

  \item \textbf{Rank elasticity to axis perturbation.}
    A $\pm 10\%$ axis-level perturbation exercise classifies 13
    countries as rank-elastic and 14 as rank-locked, with sensitivity
    concentrated in the compressed mid-range (\cref{sec:rank-elasticity}).

  \item \textbf{Axis contribution to rank differentiation.}
    Bivariate $R^2$ values, partial correlations, and a six-axis
    multiple regression ($R^2 = 0.968$) quantify each axis's marginal
    contribution to the composite (\cref{sec:axis-rank}).

  \item \textbf{Classification compression diagnostics.}
    The $\eta^2$ statistic (0.483) for the four-tier classification
    is compared with tercile and quartile alternatives
    (\cref{sec:compression-diagnostics}).

  \item \textbf{Defence axis structural analysis.}
    Sixteen of 27 member states (59.3\%) have defence as their
    dominant axis; the mean gap between the defence score and the
    next-highest axis is 0.184 (\cref{sec:defence-structural}).

  \item \textbf{Defence-dominance stress test.}
    Three counterfactual composites (without defence, without
    logistics, without both) demonstrate that the ranking retains
    structural coherence ($\rho \geq 0.654$) even when the two
    highest-variance axes are entirely excluded
    (\cref{sec:defence-stress-test}).

  \item \textbf{Formal structural typology.}
    Countries are classified by dominance gap $\Delta_i$ into
    single-axis dominated, dual-axis concentrated, and balanced
    profiles, linking profile polarisation to rank volatility
    (\cref{sec:structural-typology}).

  \item \textbf{Cross-axis significance testing with multiple-testing
    correction.}
    Formal $t$-tests with Bonferroni correction identify only two
    statistically significant correlations at $\alpha = 0.05$:
    energy--technology ($r = 0.530$, $t = 3.124$) and critical
    inputs--logistics ($r = 0.549$, $t = 3.280$)
    (\cref{sec:corr-tstat}).

  \item \textbf{PCA dimensionality diagnostic with component loadings.}
    Three principal components exceed the Kaiser criterion (76.1\%
    of variance); PC1 captures goods-trade concentration, PC2
    captures defence--logistics, and PC3 isolates the financial axis
    (\cref{tab:pca-loadings}).

  \item \textbf{Compensability analysis.}
    The max--composite gap, max--min range, and coefficient of
    variation identify the Netherlands (gap = 0.624, CV = 0.880)
    as the most internally compensated country, revealing hidden
    single-axis vulnerabilities masked by the composite
    (\cref{sec:compensability}).

  \item \textbf{Geographic and structural correlates.}
    Island status, population size, and founding-member status
    correlate with the composite ISI but not with the financial
    axis, reinforcing its orthogonality to goods-trade determinants
    (\cref{sec:geo-correlates}).

  \item \textbf{Monte Carlo and winsorization robustness.}
    Dirichlet weight perturbation (mean $\rho = 0.979$) and 95th
    percentile winsorization ($\rho = 1.000$) confirm overall ranking
    stability (\cref{sec:dirichlet,sec:winsorization}).

  \item \textbf{Eigenvalue and dimensionality analysis.}
    Three principal components exceed the Kaiser criterion, explaining
    76.1\% of variance, with an effective dimensionality of
    approximately 4.4 (\cref{tab:eigenvalues}).
\end{enumerate}

\subsection{Policy implications}

The five structural conclusions of the discussion
(\cref{sec:discussion}) carry direct policy implications:

\begin{itemize}[leftmargin=2em]
  \item The dominance of defence and logistics variance implies that
    EU strategic autonomy policy should prioritise diversification of
    defence procurement and logistics corridors over financial or
    technology import diversification, at least from the perspective
    of reducing cross-country composite variation.

  \item The mid-tier ranking fragility implies that country rankings
    in the 0.30--0.40 ISI range should be interpreted with caution
    and presented with confidence intervals or rank ranges rather than
    point estimates.

  \item The compensability findings argue for policy dashboards that
    display axis-level profiles alongside the composite, to avoid
    masking single-axis vulnerabilities.

  \item The orthogonality of the financial axis indicates that
    financial-service diversification and goods-trade diversification
    can be pursued as independent policy tracks.
\end{itemize}

\subsection{Future research}

Three directions merit further investigation:
\begin{enumerate}[leftmargin=2em]
  \item \textbf{Panel extension.}  Computing the ISI for multiple
    years would enable analysis of temporal trends, policy
    responsiveness, and the impact of specific geopolitical events
    (e.g., Russia--Ukraine conflict) on concentration dynamics.

  \item \textbf{Axis-specific deep dives.}  Each of the six axes
    could be decomposed into sub-components (e.g., the defence axis
    into military equipment categories) to provide finer-grained
    policy guidance.

  \item \textbf{Network analysis.}  Bilateral trade-flow data
    underlying the HHI computations could be used to construct
    supplier-country networks, enabling identification of common
    single-supplier dependencies across multiple EU member states.
\end{enumerate}
