% =============================================================================
% Section 1: Introduction & Contribution
% =============================================================================
\section{Introduction and Contribution}\label{sec:introduction}

\subsection{Motivation}

Composite indicators that measure structural features of national
economies have a long tradition in institutional research.  Indices
published by the OECD, the World Economic Forum, and various United
Nations agencies compress multidimensional data into scalar summaries
that facilitate cross-country comparison and policy
monitoring~\cite{oecd_handbook_2008}.  The International Sovereignty
Index (ISI) contributes to this tradition by focusing specifically on
\emph{concentration of bilateral external dependencies}: the degree to
which a country's cross-border economic and strategic linkages are
concentrated among a small number of counterparty states.

The ISI does not measure sovereignty in a political or legal sense.
Rather, it quantifies a structural feature---counterparty
concentration---that is a prerequisite for informed policy analysis of
external dependence.  A high ISI score indicates that a country's
external flows in a given domain are heavily concentrated among few
partners; a low score indicates dispersion across many partners.  The
index makes no normative claim about whether concentration is
advantageous or disadvantageous.

\subsection{Scope of this paper}

This paper---the second in the ISI series---presents the full
empirical results of the first ISI application: 27~EU member states,
vintage-2024 data, methodology v1.0.  The methodological foundations
(axis definitions, normalisation procedures, data sources, aggregation
rule, integrity chain) are documented in Paper~1~\cite{isi_paper1}.
The present paper is self-contained for readers interested in the
empirical findings; cross-references to Paper~1 are provided where
construction details are relevant to interpretation.

The specific contributions of this paper are:

\begin{enumerate}[leftmargin=2em]
  \item A complete composite and axis-level ranking of all 27~EU
    member states with eight-decimal precision, accompanied by
    classification labels.
  \item A distributional analysis of the composite index, including
    summary statistics, percentiles, a Lorenz curve, and a Gini
    coefficient.
  \item An axis-level diagnostic framework that identifies which axes
    drive cross-country variation, using a covariance-based variance
    decomposition and axis-dominance analysis.
  \item A cross-axis correlation analysis identifying moderate
    pairwise associations and representing them in a network graph.
  \item A four-part robustness programme: leave-one-axis-out (LOO)
    analysis, z-score standardisation, Dirichlet weight perturbation
    (Monte Carlo), and winsorisation of high-variance axes.
  \item Radar-chart profiles for eight selected countries illustrating
    the heterogeneity of axis-level patterns.
  \item A reproducibility protocol and computational appendix enabling
    full replication from the provided snapshot.
\end{enumerate}

\subsection{Interpretation conventions}

Throughout this paper, the following conventions apply:

\begin{itemize}[leftmargin=2em]
  \item Higher ISI composite $\Rightarrow$ higher concentration
    $\Rightarrow$ greater structural dependence on a small set of
    counterparties.
  \item Rank~1 = most concentrated country; rank~27 = least
    concentrated.
  \item The terms ``most concentrated'' and ``least concentrated'' are
    descriptive.  They are not synonymous with ``worst'' and ``best.''
  \item Correlation between axes does not imply causation.  Where
    plausible structural interpretations exist, they are stated as
    hypotheses, not conclusions.
  \item Population standard deviation ($N$ denominator) is used
    throughout because the 27~EU member states constitute the full
    target population, not a sample drawn from a larger universe.
\end{itemize}

\subsection{Structure}

The remainder of this paper is organised as follows.
\Cref{sec:data-construction} summarises the data sources, aggregation
rule, and classification thresholds.
\Cref{sec:composite-results} presents the composite distribution,
full ranking, and tier-compression analysis.
\Cref{sec:axis-diagnostics} provides axis-level diagnostics including
dispersion, dominance, and variance decomposition.
\Cref{sec:cross-axis} examines the pairwise correlation structure.
\Cref{sec:robustness} reports four robustness and sensitivity
exercises.
\Cref{sec:country-profiles} presents radar-chart profiles for eight
selected countries.
\Cref{sec:discussion} discusses interpretation boundaries and
limitations.
\Cref{sec:conclusion} concludes.
Appendices~A--D provide full tables, robustness details, the
reproducibility protocol, and the computational code listing.
