% =============================================================================
% Appendix B: Reproducibility Protocol
% =============================================================================
\section{Reproducibility Protocol}\label{app:reproducibility}

This appendix documents the procedure for reproducing all results
presented in this paper.

\subsection{Software requirements}

\begin{itemize}[leftmargin=2em]
  \item \textbf{Python} $\geq$ 3.10 (tested with 3.14.0).
  \item \textbf{NumPy} $\geq$ 2.0 (tested with 2.2.6).
    NumPy is the only external dependency.
  \item \textbf{LaTeX} distribution with pdfLaTeX (tested with
    TeX~Live 2025), including packages: \texttt{pgfplots} ($\geq$
    1.18), \texttt{tikz}, \texttt{siunitx}, \texttt{longtable},
    \texttt{booktabs}, \texttt{natbib}, \texttt{cleveref},
    \texttt{threeparttable}, \texttt{tcolorbox}, \texttt{listings},
    \texttt{xcolor}.
\end{itemize}

\subsection{Directory structure}

The project is organised as follows:

\begin{verbatim}
paper3/
+-- main.tex                 % Master document
+-- references.bib           % Bibliography
+-- isi-series.sty -> ../isi-series.sty
+-- sections/
|   +-- 00_executive_summary.tex
|   +-- 01_thesis_and_context.tex
|   +-- 02_core_finding_defence_logistics.tex
|   +-- 03_mechanics_and_proof_sketch.tex
|   +-- 04_empirics_tables_figures.tex
|   +-- 05_robustness_short.tex
|   +-- 06_implications.tex
|   +-- 07_limitations.tex
|   +-- 08_conclusion.tex
+-- tables/                  % Generated .tex table fragments
|   +-- T01_key_summary.tex
|   +-- T02_variance_decomposition.tex
|   +-- T03_axis_dominance_counts.tex
|   +-- T04_loo_rank_disruption.tex
|   +-- T05_dirichlet_volatility_top.tex
|   +-- TA1_full_variance_detail.tex
|   +-- TA2_dominance_per_country.tex
|   +-- TA3_loo_rank_diffs.tex
|   +-- F01_variance_bar.dat
|   +-- F02_dominance_share.dat
|   +-- F03_loo_rho.dat
+-- figures/                 % TikZ figure source files
|   +-- F01_variance_bar.tex
|   +-- F02_dominance_share.tex
|   +-- F03_loo_rho.tex
+-- appendix/
|   +-- appendix_a_supporting_tables.tex
|   +-- appendix_b_reproducibility.tex
|   +-- appendix_c_code_listing.tex
+-- scripts/
    +-- isi_results_paper3.py
\end{verbatim}

\subsection{Step-by-step reproduction}

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Install Python and NumPy.}
    \begin{verbatim}
    python -m venv .venv
    source .venv/bin/activate
    pip install numpy
    \end{verbatim}

  \item \textbf{Run the computation script.}
    \begin{verbatim}
    cd paper3/scripts
    python isi_results_paper3.py
    \end{verbatim}
    This generates all \texttt{.tex} table fragments and
    \texttt{.dat} coordinate files in the \texttt{tables/} directory.
    The script embeds the full {\ISI} v1.0 data snapshot and performs
    all computations (variance decomposition, axis dominance, LOO,
    Dirichlet Monte Carlo, correlation structure).

  \item \textbf{Compile the LaTeX document.}
    \begin{verbatim}
    cd paper3
    pdflatex main.tex
    bibtex main
    pdflatex main.tex
    pdflatex main.tex
    \end{verbatim}
    Three passes are required to resolve cross-references and the
    bibliography.

  \item \textbf{Verify.}  The final PDF should compile with
    0~errors, 0~warnings, and 0~overfull boxes.
\end{enumerate}

\subsection{Random seed}

The Dirichlet Monte Carlo exercise uses a fixed random seed
(\texttt{np.random.seed(42)}) to ensure exact reproducibility of
rank volatility statistics.  Changing the seed will produce
quantitatively similar but not identical results.

\subsection{Data integrity}

\begin{sloppypar}
The {\ISI} v1.0 data snapshot is embedded directly in the Python
script (\texttt{isi\_results\_\allowbreak paper3.py},
lines~42--82) as a list of tuples.
No external data files need to be downloaded or configured.  The
embedded data constitutes the single authoritative source for all
computations in this paper and is identical to the snapshot used in
Paper~2~\cite{isi_paper2}.
\end{sloppypar}
