% sections/00_abstract.tex
\begin{abstract}
\noindent
This document is the technical specification of the International Sovereignty
Index (\ISI), a composite indicator that measures the degree to which a
sovereign state concentrates its critical external dependencies on a narrow set
of counterparties.  The \ISI applies a Herfindahl--Hirschman framework across
six thematic axes---financial sovereignty, energy dependency, technology and
semiconductor dependency, defence industrial dependency, critical inputs and
raw materials dependency, and logistics and freight dependency---each
decomposed into one or two measurement channels.  Country-level concentration
scores are computed on the unit interval, arithmetically averaged into a
single composite, and mapped to a four-tier classification (\emph{highly
concentrated}, \emph{moderately concentrated}, \emph{mildly concentrated},
\emph{unconcentrated}).  A deterministic scenario engine permits first-order
sensitivity analysis by applying bounded multiplicative adjustments at the
axis-score level.  All computational outputs are bound to a cryptographic
integrity chain (SHA-256 hashing, Ed25519 signing) that enables third-party
audit without access to the computation environment.

Beyond the core methodology, this specification formalises the determinism
guarantees of the computation pipeline (ten invariants, floating-point
assumptions, canonical serialisation rules, and prohibited operations), the
snapshot materialisation architecture (directory structure, output schemas,
methodology registry, and immutability contract), the read-only API contract
(endpoint taxonomy, caching semantics, error codes, and backward-compatibility
rules), the verification and test-coverage requirements (unit tests,
property-based tests, golden-file tests, integration tests, and
integrity-chain tests), a threat model with six adversary classes and four
security goals, and a formal computational state model defining the
ten-state pipeline with explicit irreversibility gates and mutation
prohibitions.

\ISI v\,1.0 covers the EU-27.  Extension to additional country sets requires
source-level availability, harmonisation, and integrity validation.  All
scores are rounded to eight decimal places (round-half-to-even) prior to
classification and hashing.
\end{abstract}
