% sections/01_introduction.tex
\section{Introduction}\label{sec:introduction}

\subsection{Scope}\label{sec:intro:scope}

The International Sovereignty Index (\ISI) quantifies the structural
concentration of a country's critical external dependencies.  It does not
measure sovereignty in the normative or legal sense.  It measures the extent to
which a state relies on a small number of foreign partners for essential goods,
services, and capabilities across six domains: financial market access, energy
supply, technology and semiconductor inputs, defence-industrial procurement,
critical raw materials, and logistics and freight infrastructure.

The index operates at the country--year level.  Each observation produces a
composite score on the unit interval \([0,1]\) and a categorical
classification.  Higher values denote greater concentration and, by
implication, greater structural exposure to partner-specific disruption.

\ISI v\,1.0 covers the EU-27.  The computation pipeline, integrity chain,
and classification logic are country-set agnostic.  Extension to non-EU
countries is conditional on source-level data availability, cross-source
harmonisation, and integrity validation for each additional jurisdiction.

\subsection{Context}\label{sec:intro:context}

Existing composite indicators in the sovereignty-adjacent space---the
\textcite{kof2023globalisation} Globalisation Index, the
\textcite{unctad2023productive} Productive Capacities Index, and the
EU~Global Gateway scorecards---measure either \emph{openness} or
\emph{capacity}.  None directly measures \emph{concentration} of external
dependence.  The \ISI occupies a distinct methodological position: it
quantifies the structural geometry of dependence, complementing---not
substituting---openness and capacity measures.

\subsection{Design Principles}\label{sec:intro:principles}

Five principles constrain the index architecture:

\begin{enumerate}[label=\textbf{P\arabic*}.]
  \item \textbf{Transparency.}  Every formula, data source, and threshold is
    published.  No proprietary model or hidden weighting layer exists.
  \item \textbf{Reproducibility.}  Given identical input data, any competent
    third party must be able to reproduce the published scores to full
    numerical precision.
  \item \textbf{Auditability.}  Cryptographic integrity records (hash chains
    and digital signatures) accompany every published snapshot, enabling
    verification without access to the original computation environment.
  \item \textbf{Parsimony.}  The index uses exactly one concentration
    measure, the Herfindahl--Hirschman Index, applied uniformly across all
    axes.  No axis receives preferential weighting in the composite.
  \item \textbf{Conservatism.}  Where methodological choices are ambiguous,
    the index defaults to the option that is simpler, more established, and
    easier to audit.
\end{enumerate}

\subsection{Document Map}\label{sec:intro:map}

The remainder of this document is organised as follows.
\Cref{sec:data-architecture} describes the data sources, ingestion protocol,
and quality controls.  \Cref{sec:math-framework} presents the mathematical
framework---the concentration formula, rounding rules, and classification
thresholds.  \Cref{sec:axis-specifications} specifies each of the six thematic
axes.  \Cref{sec:aggregation} details the composite aggregation and ranking
procedure.  \Cref{sec:scenario-engine} documents the scenario simulation
model.  \Cref{sec:integrity-chain} describes the cryptographic integrity
mechanism.  \Cref{sec:limitations} enumerates the structural limitations of the
index.  \Cref{sec:governance} sets out the governance and versioning protocol.
\Cref{sec:determinism} formalises the determinism guarantees and computational
invariants of the pipeline.  \Cref{sec:snapshot-architecture} specifies the
snapshot materialisation architecture, output schemas, and immutability
contract.  \Cref{sec:api-contract} defines the read-only API contract,
endpoint taxonomy, and backward-compatibility rules.
\Cref{sec:verification} documents the verification test suite and coverage
requirements.  \Cref{sec:threat-model} defines the threat model, adversary
classes, security goals, trust anchors, and key-rotation policy.
\Cref{sec:state-model} presents a formal computational state model that
specifies the ten pipeline states, nine transition functions, two
irreversibility gates, and the scenario engine's isolation as a
non-state-altering branch.  \Cref{sec:architecture-overview} provides a
concise four-layer system abstraction summarising the mathematical,
computational, integrity, and governance layers.
Three reference appendices follow.
