% sections/02_data_architecture.tex
\section{Data Architecture}\label{sec:data-architecture}

\subsection{Source Registry}\label{sec:data:sources}

The \ISI draws on five primary statistical sources, each mapped to one or
more thematic axes.  \Cref{tab:source-registry} summarises the registry.

\begin{table}[htbp]
  \centering\small
  \caption{Primary data sources (v\,1.0).}\label{tab:source-registry}
  \begin{tabularx}{\textwidth}{@{}l l l X@{}}
    \toprule
    \textbf{Axis} & \textbf{Source} & \textbf{Granularity}
      & \textbf{Window} \\
    \midrule
    1\,---\,Financial   & BIS LBS / IMF CPIS
      & Counterparty, quarterly   & 2024-Q4 \\
    2\,---\,Energy      & Eurostat \texttt{nrg\_ti\_*}
      & Fuel, partner, annual     & 2024 \\
    3\,---\,Tech./Semi. & Eurostat Comext ds-045409
      & CN8, bilateral, annual    & 2023--2024 \\
    4\,---\,Defence     & SIPRI ATD (TIV)
      & System, bilateral, annual & 2019--2024 \\
    5\,---\,Crit.\ Inputs & Eurostat Comext CN8
      & CN8, bilateral, annual    & 2023--2024 \\
    6\,---\,Logistics   & Eurostat transport stats
      & Mode, partner, annual     & 2022--2024 \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Coverage}\label{sec:data:coverage}

The unit of observation is the country--year pair \((i, t)\).  The reference
population for \ISI v\,1.0 is the EU-27.  The architecture is international by
design: the computation pipeline, integrity chain, and classification logic are
country-set agnostic.  Extension to non-EU countries requires source-level data
availability, harmonisation, and integrity validation for each additional
jurisdiction.

For a given vintage year~\(t\), a country~\(i\) is included in the published
snapshot if and only if at least four of the six axes yield a computable
concentration score.  Where fewer than four axes are computable, the country is
flagged with warning code~\texttt{W-COV} and excluded from the ranked output
(see \cref{sec:appendix:warnings}).

\subsection{Data Availability Constraints (v\,1.0)}\label{sec:data:constraints}

Only \ISI~2024 is materialisable under v\,1.0.  The binding constraints are:

\begin{itemize}
  \item \textbf{BIS LBS.}  The Locational Banking Statistics are published
    with a two-quarter lag.  Only the 2024-Q4 release provides sufficient
    counterparty coverage for the EU-27.  Earlier quarters lack the partner
    disaggregation required for Channel~A of Axis~1.
  \item \textbf{Eurostat energy.}  The \texttt{nrg\_ti\_*} tables report
    annual fuel-level import data.  The 2024 vintage is the most recent
    complete release.
  \item \textbf{SIPRI.}  The Arms Transfers Database reports calendar-year
    deliveries.  The six-year window (2019--2024) is the minimum required
    to smooth lumpy transfer patterns for EU member states.
  \item \textbf{Eurostat transport.}  Modal freight statistics lag by one to
    two years.  The 2022--2024 window represents the best available coverage
    at time of computation.
\end{itemize}

An \ISI~2023 vintage is infeasible: BIS counterparty data for 2023-Q4 lack
sufficient EU-27 partner coverage, and the SIPRI window would contract below
the minimum smoothing threshold for several member states.  Time-series
extension requires additional source vintages and is not supported in v\,1.0.

\subsection{Ingestion Protocol}\label{sec:data:ingestion}

Data ingestion follows a four-stage pipeline:

\begin{enumerate}
  \item \textbf{Acquisition.}  Raw files are retrieved from the source API or
    bulk-download portal in the format published by the provider (CSV, SDMX,
    JSON).
  \item \textbf{Parsing.}  Source-specific parsers normalise records into a
    common schema per axis.  Each parser is a dedicated module in the backend
    pipeline.
  \item \textbf{Validation.}  Row-level checks enforce non-negative values,
    valid ISO~3166-1 alpha-2 codes (EU member states), and year-range bounds.
    Records that fail validation are quarantined and logged.
  \item \textbf{Loading.}  Validated records are loaded into the computation
    layer.  No post-load transformation occurs outside the formula
    definitions in \cref{sec:math-framework}.
\end{enumerate}

\subsection{Quality Controls}\label{sec:data:quality}

Three quality controls apply at the axis level:

\begin{description}[style=nextline]
  \item[Temporal gap detection.]
    If the most recent available data point for axis~\(a\) and country~\(i\)
    predates the snapshot vintage year~\(t\) by more than two years, the axis
    score is computed from the latest available year and warning
    code~\texttt{W-LAG} is attached.
  \item[Partner-set completeness.]
    If the share vector for axis~\(a\) and country~\(i\) covers fewer than
    \SI{90}{\percent} of the total reported flow (due to missing partner
    disaggregation), warning code~\texttt{W-PAR} is attached.
  \item[Channel fallback basis.]
    Where an axis defines two channels but only one is computable for a given
    country, the axis score reverts to the surviving channel.  The fallback
    basis (\texttt{BOTH}, \texttt{A\_ONLY}, \texttt{B\_ONLY}, \texttt{NONE})
    is recorded in the output metadata.
\end{description}

Warning codes are propagated to the published output but do not alter the
numerical score.  They serve as metadata for downstream consumers.