% sections/07_integrity_chain.tex
\section{Integrity Chain}\label{sec:integrity-chain}

\subsection{Design Rationale}\label{sec:integrity:rationale}

Published index scores are only credible if consumers can verify that the
scores have not been altered after computation.  The \ISI embeds a
cryptographic integrity chain in every published snapshot.  The chain provides
three guarantees:

\begin{enumerate}
  \item \textbf{Tamper evidence.}  Any post-publication modification to a
    country score, axis score, weight, threshold, or classification will
    invalidate the hash chain.
  \item \textbf{Authorship attestation.}  The Ed25519 signature over the
    snapshot hash binds the publication to a specific signing key, allowing
    consumers to verify that the snapshot was produced by the authorised
    computation environment.
  \item \textbf{Offline verification.}  Verification requires only the
    published snapshot files, the public key, and a standard cryptographic
    library.  No access to the computation environment or raw data is needed.
\end{enumerate}

\subsection{Country-Level Hash}\label{sec:integrity:country}

For each country~\(i\) in vintage year~\(t\), a canonical string
representation is constructed by concatenating the following fields in
fixed order:

\begin{enumerate}
  \item ISO~3166-1 alpha-2 country code.
  \item Vintage year.
  \item Methodology version string (e.g., \texttt{1.0}).
  \item Six axis scores \(M_i^{(1)}, \ldots, M_i^{(6)}\), each represented
    as a fixed-point decimal with eight digits after the decimal point.
    Null axes are encoded as the string \texttt{null}.
  \item Composite score (eight-decimal fixed-point representation).
  \item Weight vector (the axis weights used; \(\tfrac{1}{6}\) for all six
    axes in v\,1.0).
  \item Classification thresholds (the four boundary values).
\end{enumerate}

The country-level hash is:
%
\[
  h_i \;=\; \mathrm{SHA\text{-}256}(\text{canonical}_i)
\]

The exact serialisation format (field separators, encoding) is defined in
the backend module \texttt{snapshot\_integrity.py} and is frozen under
major version~1.x.  Any change to the serialisation order or encoding
constitutes a major-version increment.

\subsection{Snapshot Hash}\label{sec:integrity:snapshot}

The snapshot hash aggregates all country-level hashes into a single digest.
Country hashes are sorted in ascending lexicographic order of country code
and concatenated:
%
\[
  H_t \;=\; \mathrm{SHA\text{-}256}\!\bigl(h_{\mathrm{AT}} \;\|\;
  h_{\mathrm{BE}} \;\|\; \cdots \;\|\; h_{\mathrm{SK}}\bigr)
\]
%
where \(\|\) denotes concatenation.

\subsection{Digital Signature}\label{sec:integrity:signature}

The snapshot hash is signed using Ed25519 \parencite{bernstein2012eddsa}:
%
\[
  \sigma_t \;=\; \mathrm{Ed25519\text{-}Sign}(sk,\; H_t)
\]
%
where \(sk\) is the signing private key held by the publication authority.
The corresponding public key~\(pk\) is distributed via the project's public
key registry and pinned in the snapshot metadata.

\subsection{Published Artefacts}\label{sec:integrity:artefacts}

Each snapshot publication includes three integrity files:

\begin{description}[style=nextline]
  \item[\texttt{MANIFEST.json}]
    Lists every output file in the snapshot (score files, metadata,
    methodology registry) with its SHA-256 hash.  Consumers can verify
    file-level integrity by recomputing hashes and comparing.
  \item[\texttt{HASH\_SUMMARY.json}]
    Contains the per-country hashes~\(h_i\) and the aggregate snapshot
    hash~\(H_t\).  This file enables country-level verification without
    recomputing the full pipeline.
  \item[\texttt{SIGNATURE.json}]
    Contains the Ed25519 signature~\(\sigma_t\) over~\(H_t\) and the
    public key identifier used for signing.
\end{description}

\subsection{Verification Procedure}\label{sec:integrity:verify}

A consumer verifying snapshot~\(t\) performs the following steps:

\begin{enumerate}
  \item Recompute each country hash~\(h_i\) from the published score record
    using the canonical serialisation defined in
    \texttt{snapshot\_integrity.py}.
  \item Concatenate country hashes in lexicographic order and compute
    \(H_t'\).
  \item Verify that \(H_t' = H_t\) (the published snapshot hash in
    \texttt{HASH\_SUMMARY.json}).
  \item Verify the Ed25519 signature:
    \(\mathrm{Ed25519\text{-}Verify}(pk, H_t, \sigma_t) = \text{true}\)
    using the public key referenced in \texttt{SIGNATURE.json}.
\end{enumerate}

If any step fails, the snapshot is considered compromised.

\subsection{Key Management}\label{sec:integrity:keys}

The signing key pair is generated once per major version of the index.  Key
rotation occurs at major-version boundaries (see \cref{sec:governance}).  The
public key is published in the project repository and in
\texttt{SIGNATURE.json}.  The private key is stored in an isolated
environment and is never exposed to the general computation pipeline.