% sections/10_determinism.tex
\section{Determinism and Computational Guarantees}\label{sec:determinism}

This section formalises the determinism properties of the \ISI computation
pipeline.  Every guarantee stated here is a binding contract: a conforming
implementation \emph{must} satisfy all invariants, and a failure to satisfy
any single invariant constitutes a non-conforming build.

\subsection{Scope of Determinism}\label{sec:det:scope}

The \ISI pipeline is \textbf{fully deterministic}: given identical input data,
identical methodology-version parameters, and a conforming arithmetic
environment, the pipeline produces bit-identical output across invocations,
platforms, and calendar time.  Determinism is not a best-effort aspiration; it
is an architectural invariant enforced by the design constraints enumerated
below.

Determinism extends to all published artefacts: axis scores, composite scores,
classifications, rankings, warning codes, hash digests, and the snapshot
signature.  It does \emph{not} extend to execution time, memory consumption,
or log output, which are platform-dependent observables.

\subsection{Formal Invariants}\label{sec:det:invariants}

The following invariants hold for every conforming implementation of the
\ISI v\,1.0 pipeline.  Each invariant is designated \textbf{D-\textit{n}} and
is referenced in the test-suite specification (\cref{sec:verification}).

\begin{description}[style=nextline, leftmargin=3.5em, labelwidth=3em]

  \item[\textbf{D-1}] \textbf{Input--output identity.}\enspace
    Let \(\mathcal{D}\) be a fixed input dataset and \(\theta\) a fixed
    methodology-parameter vector.  For any two invocations \(r_1, r_2\) of
    the pipeline on \((\mathcal{D}, \theta)\):
    \[
      \mathrm{output}(r_1) \;=\; \mathrm{output}(r_2)
    \]
    where equality is byte-level identity of the serialised output.

  \item[\textbf{D-2}] \textbf{Platform invariance.}\enspace
    Let \(P_1, P_2\) be any two platforms satisfying the arithmetic
    assumptions of \cref{sec:det:floating-point}.  Then:
    \[
      \mathrm{output}(P_1, \mathcal{D}, \theta)
      \;=\;
      \mathrm{output}(P_2, \mathcal{D}, \theta)
    \]

  \item[\textbf{D-3}] \textbf{Time invariance.}\enspace
    The pipeline output depends only on \((\mathcal{D}, \theta)\).  No
    timestamp, random seed, environment variable, locale setting, or
    system clock reading enters the computation.  Formally: for any
    calendar times \(t_1, t_2\):
    \[
      \mathrm{output}(t_1, \mathcal{D}, \theta)
      \;=\;
      \mathrm{output}(t_2, \mathcal{D}, \theta)
    \]

  \item[\textbf{D-4}] \textbf{Order invariance of input rows.}\enspace
    The pipeline sorts all input records into a canonical order before
    computation.  Permuting the row order of the input dataset produces
    identical output:
    \[
      \forall \pi \in \mathcal{S}_{|\mathcal{D}|}:\quad
      \mathrm{output}(\pi(\mathcal{D}), \theta)
      \;=\;
      \mathrm{output}(\mathcal{D}, \theta)
    \]

  \item[\textbf{D-5}] \textbf{Rounding closure.}\enspace
    All published scores are rounded to eight decimal places
    (round-half-to-even) \emph{before} any downstream consumption
    (classification, hashing, serialisation).  Formally: let \(R_8(x)\)
    denote the round-half-to-even function at precision~8.  Then for every
    published score~\(s\):
    \[
      R_8(s) \;=\; s
    \]
    That is, every published score is a fixed point of the rounding operator.

  \item[\textbf{D-6}] \textbf{Hash determinism.}\enspace
    The country-level hash (\cref{sec:integrity:country}) and snapshot hash
    (\cref{sec:integrity:snapshot}) are fully determined by the published
    scores and the canonical serialisation format.  No non-deterministic
    input (timestamp, hostname, process identifier) enters the hash
    preimage.

  \item[\textbf{D-7}] \textbf{Classification determinism.}\enspace
    Classification is a pure function of the rounded composite score and
    the threshold vector.  For a given rounded composite~\(s\) and threshold
    vector~\(\tau = (0.15,\; 0.25,\; 0.50)\):
    \[
      \mathrm{classify}(s, \tau) \;=\;
      \begin{cases}
        \texttt{highly\_concentrated}     & \text{if } s \geq 0.50 \\
        \texttt{moderately\_concentrated} & \text{if } 0.25 \leq s < 0.50 \\
        \texttt{mildly\_concentrated}     & \text{if } 0.15 \leq s < 0.25 \\
        \texttt{unconcentrated}           & \text{if } s < 0.15
      \end{cases}
    \]
    No stochastic or environment-dependent branching is permitted.

  \item[\textbf{D-8}] \textbf{Ranking determinism.}\enspace
    The ranking function is a deterministic total order.  The primary sort
    key is the rounded composite score (descending).  The secondary sort key
    (tie-breaker) is the ISO~3166-1 alpha-2 code (ascending,
    lexicographic).  This total order is unique for any distinct set of
    (score, code) pairs.

  \item[\textbf{D-9}] \textbf{Scenario determinism.}\enspace
    The scenario engine (\cref{sec:scenario-engine}) is a deterministic
    function of the baseline scores and the adjustment vector
    \((\alpha_1, \ldots, \alpha_6)\).  Clamping (\cref{eq:scenario}) is
    deterministic; no randomised rounding or stochastic perturbation is
    applied.

  \item[\textbf{D-10}] \textbf{Missing-axis determinism.}\enspace
    The missing-axis protocol (\cref{sec:agg:missing}) is a deterministic
    function of the set of computable axes \(\mathcal{A}_i\).  The
    composite denominator is \(|\mathcal{A}_i|\), not a stochastic or
    heuristic estimate.

\end{description}

\subsection{Floating-Point Assumptions}\label{sec:det:floating-point}

The pipeline assumes \textbf{IEEE~754 binary64} (double-precision)
floating-point arithmetic with the following constraints:

\begin{enumerate}
  \item \textbf{Rounding mode.}  The default IEEE~754 rounding mode
    (\emph{round-to-nearest, ties-to-even}) must be active throughout the
    computation.  No midstream rounding-mode changes are permitted.

  \item \textbf{Operation order.}  Summations over partner shares and over
    axis scores follow a fixed left-to-right accumulation order after
    sorting the operands into canonical sequence (partner code ascending for
    share vectors; axis index ascending for composite averaging).  The
    accumulation order is specified to eliminate platform-dependent
    reordering that could produce different floating-point round-off
    patterns.

  \item \textbf{No fused multiply-add.}  Fused multiply-add (FMA)
    instructions may produce different round-off behaviour from separate
    multiply-then-add sequences.  Conforming implementations must disable
    FMA contraction or demonstrate bit-identical results under FMA.

  \item \textbf{Decimal rounding.}  The final rounding step
    (\texttt{ROUND\_PRECISION\,=\,8}, round-half-to-even) is performed
    using Python's \texttt{decimal.Decimal} module or an equivalent
    arbitrary-precision decimal library, not by native binary64 rounding.
    This eliminates the double-rounding problem inherent in converting
    binary64 to decimal.

  \item \textbf{No extended precision.}  Intermediate calculations must not
    use x87 80-bit extended precision.  All intermediates are stored and
    operated upon as binary64.  On x86 platforms, the compiler must be
    configured to use SSE2 instructions (not x87 FPU) for floating-point
    arithmetic, or the FPU control word must be set to 64-bit precision.
\end{enumerate}

These constraints are sufficient to guarantee \textbf{D-1} and \textbf{D-2}
across all platforms that conform to IEEE~754-2008.

\subsection{Canonical Serialisation}\label{sec:det:serialisation}

Deterministic output requires a deterministic serialisation format.  The
\ISI pipeline serialises all published data according to the following
normative rules.  Violation of any rule produces a non-conforming snapshot
whose hash MUST differ from that of a conforming snapshot.

\begin{enumerate}
  \item \textbf{Character encoding.}  All output files SHALL use UTF-8
    without byte-order mark (BOM).  No other encoding is permitted.

  \item \textbf{Newline convention.}  Line endings SHALL be Unix-style
    (\texttt{LF}, U+000A).  No carriage-return characters (\texttt{CR},
    U+000D) SHALL appear in any output file.

  \item \textbf{No trailing whitespace.}  No line in any output file
    SHALL end with a space (U+0020) or horizontal tab (U+0009) character
    before the line-feed terminator.

  \item \textbf{JSON formatting.}  JSON output files SHALL use two-space
    indentation, no trailing commas, and keys sorted in \emph{ascending
    lexicographic order} (byte-wise comparison of UTF-8 code units) at
    every nesting level.  Floating-point values SHALL be serialised as
    fixed-point decimal strings with exactly eight digits after the
    decimal point and a period (U+002E) as the decimal separator
    (e.g., \texttt{"0.34218753"}, not \texttt{0.34218753} or
    \texttt{"3.4218753e-01"}).  No locale-dependent thousands separator,
    comma-as-decimal, or scientific notation is permitted.

  \item \textbf{CSV formatting.}  CSV output files SHALL use comma
    delimiters, double-quote text-field quoting, and header rows matching
    the schema defined in the output specification
    (\cref{sec:snapshot:output}).  Fields SHALL be ordered by the fixed
    column schema; no column reordering is permitted across versions.

  \item \textbf{Locale independence.}  No serialisation step SHALL read
    or depend on the system locale, the \texttt{LC\_NUMERIC} environment
    variable, or any locale-sensitive formatting library.  All numeric
    formatting SHALL use the explicit rules defined in items~4 and~5.

  \item \textbf{Stable dictionary ordering.}  Where an output structure is
    constructed from an in-memory dictionary or hash map, the serialisation
    step SHALL sort keys into ascending lexicographic order before writing.
    Reliance on insertion-order preservation (e.g., Python~3.7+ dictionary
    ordering) is insufficient; an explicit \texttt{sorted()} call is
    REQUIRED.

  \item \textbf{Deterministic country-code sorting.}  Country codes
    in all output files (CSV rows, JSON arrays, hash-input concatenation)
    SHALL be sorted in ascending lexicographic order of the two-character
    ISO~3166-1 alpha-2 code.

  \item \textbf{Hash-input strings.}  The canonical string used as the
    SHA-256 preimage for country-level hashing
    (\cref{sec:integrity:country}) concatenates fields using the pipe
    character (\texttt{|}) as a delimiter, in the fixed order specified
    in the integrity-chain section.  No whitespace padding is applied.
    Null axis scores are encoded as the literal string \texttt{null}.
\end{enumerate}

\paragraph{Hash-breakage clause.}
If any of the serialisation rules above is violated---even if the
underlying numeric values are correct---the resulting SHA-256 hash
SHALL differ from the hash of a conforming serialisation.  This is a
design feature, not a defect: it ensures that serialisation conformance
is verifiable through the integrity chain.

\subsection{Prohibited Operations}\label{sec:det:prohibited}

The following operations are prohibited in the computation pipeline:

\begin{itemize}
  \item \textbf{Random number generation.}  No call to a pseudo-random or
    true-random number generator is permitted at any stage.
  \item \textbf{System clock queries.}  No timestamp or system clock
    reading may influence a computation step.  Timestamps in metadata
    (e.g., the \texttt{computed\_at} field in the snapshot manifest) are
    recorded as post-computation annotations and are excluded from the
    hash preimage.
  \item \textbf{Environment-dependent branching.}  No control-flow decision
    may depend on the hostname, operating system, CPU architecture, locale,
    timezone, or any environment variable.
  \item \textbf{Unordered iteration.}  No computation may iterate over a
    hash-map, set, or dictionary without first imposing a canonical sort
    order.  In Python, this implies explicit \texttt{sorted()} calls on
    dictionary keys wherever iteration order affects the result.
  \item \textbf{Floating-point equality comparison for branching.}
    Classification uses strict inequality comparisons against decimal
    threshold constants.  No floating-point equality test
    (e.g., \texttt{==\,0.25}) is used in any branching condition, except
    where both operands are known to be exactly representable in binary64.
\end{itemize}

\subsection{Hash Dependency Graph}\label{sec:det:hash-graph}

The integrity chain imposes a strict dependency order on computed values.
No hash may be computed until all of its upstream inputs have been finalised
(rounded and serialised).  The dependency graph is:

\begin{quote}\small
\begin{enumerate}[label=\textbf{Stage~\arabic*}., leftmargin=4em]
  \item \textbf{Channel scores.}\enspace  Per-partner share vectors
    \(\to\) channel-level \HHI (\cref{eq:hhi}).
  \item \textbf{Axis scores.}\enspace  Channel scores \(\to\)
    axis-level aggregation (\cref{eq:dual-channel,eq:weighted-b,eq:fuel-avg}).
  \item \textbf{Composite score.}\enspace  Six axis scores \(\to\)
    arithmetic mean (\cref{eq:composite}).
  \item \textbf{Rounding.}\enspace  All scores rounded to 8~decimal
    places (round-half-to-even).
  \item \textbf{Classification.}\enspace  Rounded composite \(\to\)
    four-tier label (\cref{tab:thresholds}).
  \item \textbf{Country hash.}\enspace  Canonical serialisation of
    rounded scores, classification, and metadata \(\to\)
    \(h_i = \mathrm{SHA\text{-}256}(\text{canonical}_i)\)
    (\cref{sec:integrity:country}).
  \item \textbf{Snapshot hash.}\enspace  Sorted country hashes
    \(\to\) \(H_t = \mathrm{SHA\text{-}256}(h_{\mathrm{AT}} \| \cdots \|
    h_{\mathrm{SK}})\) (\cref{sec:integrity:snapshot}).
  \item \textbf{Signature.}\enspace  Snapshot hash \(\to\)
    \(\sigma_t = \mathrm{Ed25519\text{-}Sign}(sk, H_t)\)
    (\cref{sec:integrity:signature}).
\end{enumerate}
\end{quote}

A stage may commence only after all inputs from the preceding stage are
finalised.  Rounding (Stage~4) is the single irreversible precision-reduction
step; all subsequent stages (5--8) consume only rounded values.  Reversing
the order of Stages~4 and~5 (i.e., classifying before rounding) would violate
invariant \textbf{D-5} and produce inconsistent hash preimages.

\subsection{Determinism Certification}\label{sec:det:certification}

A conforming implementation is certified deterministic if it passes the
full test suite defined in \cref{sec:verification}, including:

\begin{enumerate}
  \item \textbf{Golden-file tests.}  The output for a fixed reference
    dataset matches a pre-computed golden file byte-for-byte.
  \item \textbf{Cross-platform tests.}  The same golden-file comparison
    passes on at least two distinct platforms (e.g., x86-64 Linux and
    ARM64 macOS).
  \item \textbf{Idempotency tests.}  Running the pipeline twice on the same
    input produces byte-identical output (verifying \textbf{D-1}).
  \item \textbf{Permutation tests.}  Randomly permuting the input-row order
    produces identical output (verifying \textbf{D-4}).
\end{enumerate}
