% sections/15_state_model.tex
\section{Formal Computational State Model}\label{sec:state-model}

This section defines the \ISI computation pipeline as a deterministic finite
state machine.  The model enumerates every intermediate state, every
transition function, and every irreversibility constraint.  Conforming
implementations SHALL realise the state transitions in the order specified
below; no reordering, merging, or elision of states is permitted.

% -------------------------------------------------------------------------
\subsection{State Definitions}\label{sec:state:defs}

The pipeline comprises ten discrete states, designated \textbf{S0}
through~\textbf{S9}.  Each state is a typed artefact; the transition from
state~\(n\) to state~\(n+1\) is a pure function consuming only the
artefact at state~\(n\) (and, where noted, the methodology-parameter
vector~\(\theta\)).

\begin{description}[style=nextline, leftmargin=4em, labelwidth=3.5em]

  \item[\textbf{S0}] \textbf{Raw data.}\enspace
    The unprocessed source files as received from statistical agencies.
    Format: heterogeneous (CSV, SDMX, JSON).  No computational
    transformation has been applied.  State~S0 is the sole external
    input to the pipeline.

  \item[\textbf{S1}] \textbf{Validated data.}\enspace
    Source files ingested, parsed, and validated against the quality
    controls of \cref{sec:data:quality}.  Records failing validation are
    rejected.  Output: a normalised tabular dataset with uniform column
    schema, country codes in ISO~3166-1 alpha-2, and numeric fields in
    native floating-point representation.

  \item[\textbf{S2}] \textbf{Channel concentration vectors.}\enspace
    Per-country, per-axis, per-channel partner-share vectors and the
    corresponding \HHI scores computed via \cref{eq:hhi}.  Output: a set
    of \((i, a, c, \mathrm{HHI}_{i,a,c})\) tuples for every country~\(i\),
    axis~\(a\), and channel~\(c\).

  \item[\textbf{S3}] \textbf{Axis-level scores.}\enspace
    Channel scores aggregated to axis level via the aggregation variants
    specified in \cref{sec:axis-specifications}: dual-channel averaging
    (\cref{eq:dual-channel}), category-weighted averaging
    (\cref{eq:weighted-b}), or single-channel fuel averaging
    (\cref{eq:fuel-avg}).  Output: a set of \((i, a, M_i^{(a)})\) tuples.
    Scores at this stage are in full binary64 precision.

  \item[\textbf{S4}] \textbf{Composite score.}\enspace
    The arithmetic mean of computable axis scores
    (\cref{eq:composite}).  Output: a set of \((i, \mathrm{ISI}_i)\) pairs
    in full binary64 precision.

  \item[\textbf{S5}] \textbf{Rounded and classified country object.}\enspace
    All scores (axis and composite) rounded to eight decimal places
    (round-half-to-even) using an arbitrary-precision decimal library.
    The classification function (\cref{sec:math:thresholds}) is applied to
    the rounded composite.  The ranking function (\cref{sec:agg:ranking})
    produces ordinal ranks.  Output: a fully populated country record
    containing rounded scores, classification label, rank, warning codes,
    and computable-axis count.

  \item[\textbf{S6}] \textbf{Country hash.}\enspace
    The canonical serialisation of each country record
    (\cref{sec:integrity:country}) is hashed with SHA-256 to produce
    \(h_i\).  Output: a set of \((i, h_i)\) pairs.

  \item[\textbf{S7}] \textbf{Snapshot hash.}\enspace
    Country hashes concatenated in ISO~alpha-2 lexicographic order,
    combined with the manifest hash, and hashed with SHA-256 to
    produce~\(H_t\) (\cref{sec:integrity:snapshot}).  Output: a single
    256-bit digest.

  \item[\textbf{S8}] \textbf{Signed snapshot.}\enspace
    The snapshot hash is signed with Ed25519 (\cref{sec:integrity:signature})
    to produce~\(\sigma_t\).  The snapshot directory is materialised on disk
    (\cref{sec:snapshot-architecture}).  Output: the complete snapshot
    directory including all score files, integrity files, and the signature.

  \item[\textbf{S9}] \textbf{Served API artefact.}\enspace
    The signed snapshot is loaded into the serving layer
    (\cref{sec:api-contract}).  API responses are read-only projections of
    the snapshot.  No transformation, re-computation, or re-serialisation
    occurs at serving time.

\end{description}

% -------------------------------------------------------------------------
\subsection{Transition Functions}\label{sec:state:transitions}

Each transition is designated \textbf{T\textit{n}} and maps from
state~\(\textbf{S(n{-}1)}\) to state~\(\textbf{Sn}\).  All transitions
are pure functions: they read only their input state and the
methodology-parameter vector~\(\theta\); they produce no side effects.

\begin{description}[style=nextline, leftmargin=4em, labelwidth=3.5em]

  \item[\textbf{T1}] \(f_1 : (\mathrm{S0}, \theta) \to \mathrm{S1}\).\\
    Ingestion and validation.  Applies schema enforcement, type coercion,
    country-code normalisation, and quality-control rejection rules
    (\cref{sec:data:ingestion,sec:data:quality}).

  \item[\textbf{T2}] \(f_2 : (\mathrm{S1}, \theta) \to \mathrm{S2}\).\\
    Share-vector construction and \HHI computation
    (\cref{eq:hhi}).  Partner shares are normalised to sum to~\(1.0\) per
    country--axis--channel group.  Summation follows the canonical
    accumulation order specified in \cref{sec:det:floating-point}.

  \item[\textbf{T3}] \(f_3 : (\mathrm{S2}, \theta) \to \mathrm{S3}\).\\
    Cross-channel aggregation.  Applies the axis-specific aggregation
    variant (\cref{sec:axis-specifications}).  Output scores are in full
    binary64 precision.

  \item[\textbf{T4}] \(f_4 : (\mathrm{S3}, \theta) \to \mathrm{S4}\).\\
    Composite averaging (\cref{eq:composite}).  The divisor is the number
    of computable axes~\(|\mathcal{A}_i|\), not a fixed constant.

  \item[\textbf{T5}] \(f_5 : (\mathrm{S4}, \theta) \to \mathrm{S5}\).\\
    Rounding, classification, and ranking.  This is the \textbf{rounding
    gate}: the irreversible precision-reduction step.  After this
    transition, all scores are fixed-point 8-decimal values and SHALL NOT
    revert to binary64 representation.

  \item[\textbf{T6}] \(f_6 : \mathrm{S5} \to \mathrm{S6}\).\\
    Country hashing.  Canonical serialisation of the country record
    followed by SHA-256 (\cref{sec:integrity:country}).  This is the
    \textbf{hashing gate}: once a country hash is computed, the underlying
    score values are cryptographically committed.

  \item[\textbf{T7}] \(f_7 : \mathrm{S6} \to \mathrm{S7}\).\\
    Snapshot hash construction.  Concatenation of ordered country hashes
    and manifest hash, followed by SHA-256
    (\cref{sec:integrity:snapshot}).

  \item[\textbf{T8}] \(f_8 : \mathrm{S7} \to \mathrm{S8}\).\\
    Signing and materialisation.  Ed25519 signature over~\(H_t\), followed
    by file-system materialisation of the complete snapshot directory.

  \item[\textbf{T9}] \(f_9 : \mathrm{S8} \to \mathrm{S9}\).\\
    Loading into the serving layer.  Read-only memory mapping of the
    snapshot directory.  No mutation of the snapshot is permitted during or
    after this transition.

\end{description}

% -------------------------------------------------------------------------
\subsection{Irreversibility Points}\label{sec:state:irreversibility}

Two transitions impose irreversible commitments on the pipeline state.
After each irreversibility point, the preceding full-precision
representation is discarded and SHALL NOT be recoverable from the
published artefacts.

\begin{description}[style=nextline, leftmargin=6.5em, labelwidth=6em]

  \item[\textbf{T5} (Rounding gate)]
    After transition~\textbf{T5}, all published scores exist only in
    rounded 8-decimal form.  The full binary64 intermediates are not
    serialised, not hashed, and not recoverable.  Any downstream
    computation (classification, hashing, serving) operates exclusively on
    rounded values.  Invariant \textbf{D-5} guarantees that no further
    rounding is required.

  \item[\textbf{T6} (Hashing gate)]
    After transition~\textbf{T6}, the content of each country record is
    cryptographically committed via SHA-256.  Any modification to a rounded
    score, classification label, or warning code after~\textbf{T6} would
    produce a different country hash and therefore a different snapshot hash,
    invalidating the signature.
\end{description}

The two gates partition the pipeline into three zones:

\begin{enumerate}
  \item \textbf{Pre-rounding zone} (S0--S4).  Full binary64 precision.
    Scores may be recomputed freely.
  \item \textbf{Post-rounding / pre-hashing zone} (S5).  Rounded scores
    exist but are not yet cryptographically committed.  Recomputation is
    permitted only if the rounding step is re-executed.
  \item \textbf{Post-hashing zone} (S6--S9).  Scores are committed.
    Modification is prohibited (\cref{sec:snapshot:immutability}).
\end{enumerate}

% -------------------------------------------------------------------------
\subsection{State Mutation Prohibition}\label{sec:state:mutation}

The pipeline is a strict feed-forward computation.  The following mutation
rules SHALL be enforced:

\begin{enumerate}
  \item No transition function \(f_n\) MAY modify the artefact at
    state~\(\mathrm{S(n{-}1)}\).  Transitions are pure functions that
    \emph{read} input states and \emph{produce} output states.

  \item After transition~\textbf{T8} (signing), the snapshot directory
    SHALL NOT be modified by any subsequent operation, including
    \textbf{T9} (serving).  The serving layer operates on a read-only
    view of the materialised snapshot.

  \item No transition function MAY write to a global mutable variable,
    file, or database that is read by a later transition.  All inter-state
    communication is through the explicitly typed artefacts defined in
    \cref{sec:state:defs}.
\end{enumerate}

% -------------------------------------------------------------------------
\subsection{Scenario Engine as Non-State-Altering Branch}%
\label{sec:state:scenario}

The scenario engine (\cref{sec:scenario-engine}) operates as a
\textbf{read-only branch} from state~\textbf{S5}.  It consumes the
rounded baseline scores and the user-supplied adjustment
vector~\((\alpha_1, \ldots, \alpha_6)\), applies the multiplicative
model (\cref{eq:scenario}), recomputes the composite and rank, and
returns the result to the caller.

The scenario engine SHALL NOT:

\begin{itemize}
  \item Modify any file in the snapshot directory.
  \item Write to the methodology registry.
  \item Produce or modify country hashes, snapshot hashes, or signatures.
  \item Persist scenario results in any durable store that is visible to
    the canonical pipeline.
\end{itemize}

Scenario outputs are \textbf{ephemeral}: they exist only in the API
response and are not covered by the integrity chain.  Scenario responses
SHALL NOT include a signature or hash digest.  Any downstream consumer of
scenario data MUST treat it as unsigned and non-authoritative.

Formally, the scenario engine defines a side-branch:
\[
  f_{\mathrm{scenario}} : (\mathrm{S5}, \vec{\alpha}) \to
  \mathrm{S5'_{\mathrm{ephemeral}}}
\]
where \(\mathrm{S5'_{\mathrm{ephemeral}}}\) is a transient artefact that
is never committed to~\textbf{S6} or any subsequent state.  The canonical
state path S0\,\(\to\)\,\ldots\,\(\to\)\,S9 is unaffected.

% -------------------------------------------------------------------------
\subsection{Deterministic Transition Constraints}\label{sec:state:constraints}

Every transition function \(f_n\) SHALL satisfy the following properties:

\begin{enumerate}
  \item \textbf{Purity.}\enspace  \(f_n\) depends only on its input state
    and~\(\theta\).  No hidden input (clock, locale, environment variable,
    random seed) is consumed.  This is a restatement of invariants
    \textbf{D-1} through \textbf{D-3}.

  \item \textbf{Totality.}\enspace  \(f_n\) is defined for all valid inputs.
    Partial failures (e.g., a missing axis score) are handled by the
    missing-axis protocol (\cref{sec:agg:missing}), not by undefined
    behaviour.

  \item \textbf{Convergence.}\enspace  \(f_n\) terminates in finite time
    for all valid inputs.  No unbounded iteration or recursion is permitted.

  \item \textbf{Idempotence of read operations.}\enspace  Reading from
    a state artefact does not modify it.  This is trivially satisfied by
    file-system reads but must be explicitly enforced for in-memory data
    structures that might use lazy evaluation or memoisation with side
    effects.

  \item \textbf{Canonical ordering.}\enspace  Where the transition
    iterates over a collection (partner shares, country records, axis
    scores), the iteration order SHALL be the canonical order defined in
    \cref{sec:det:floating-point} and \cref{sec:det:serialisation}.
\end{enumerate}

% -------------------------------------------------------------------------
\subsection{State Diagram Summary}\label{sec:state:diagram}

The following notation summarises the state machine.  Each arrow is labelled
with its transition function and the section that specifies it.

\begin{quote}\small
\begin{verbatim}
  S0 --[T1: ingest/validate (sec 2.4)]--> S1
  S1 --[T2: HHI computation  (sec 3.2)]--> S2
  S2 --[T3: axis aggregation (sec 3.3)]--> S3
  S3 --[T4: composite mean   (sec 3.6)]--> S4
  S4 --[T5: round+classify   (sec 3.7)]--> S5  *ROUNDING GATE*
  S5 --[T6: country hash     (sec 7.2)]--> S6  *HASHING GATE*
  S6 --[T7: snapshot hash    (sec 7.3)]--> S7
  S7 --[T8: sign+materialise (sec 7.4)]--> S8
  S8 --[T9: serve            (sec 12) ]--> S9

  S5 --[f_scenario (sec 6)]--> S5'_ephemeral  (branch)
\end{verbatim}
\end{quote}

No backward edge exists.  The graph is a directed acyclic path with a
single read-only branch at~\textbf{S5}.  This structure guarantees that
every state is reachable from~\textbf{S0} by a unique path and that no
cyclic dependency can arise.
