% =============================================================================
% Appendix B: Robustness Detail
% =============================================================================
\section{Robustness Detail Tables}\label{app:robustness}

This appendix provides the full country-level detail for each
robustness exercise summarised in \cref{sec:robustness}.

\subsection{Leave-one-axis-out: full rank table}

\Cref{tab:loo-detail} reports the rank of each country under each
LOO variant (baseline, drop axis 1, \ldots, drop axis~6), along with
the maximum rank change from baseline.

\input{tables/TB1_loo_detail}

\subsection{Individual LOO axis tables}

The following tables report the full score and rank for each country
under each individual axis omission.

\subsubsection{LOO: Financial axis omitted}
\input{tables/TC_loo_financial}

\clearpage
\subsubsection{LOO: Energy axis omitted}
\input{tables/TC_loo_energy}

\clearpage
\subsubsection{LOO: Technology axis omitted}
\input{tables/TC_loo_technology}

\clearpage
\subsubsection{LOO: Defence axis omitted}
\input{tables/TC_loo_defence}

\clearpage
\subsubsection{LOO: Critical inputs axis omitted}
\input{tables/TC_loo_critical_inputs}

\clearpage
\subsubsection{LOO: Logistics axis omitted}
\input{tables/TC_loo_logistics}

\clearpage
\subsection{Z-score standardisation: full rank comparison}

\Cref{tab:zscore-ranks} in \cref{sec:zscore} compares baseline and
z-score ranks for all~27 countries.  See \cref{tab:zscore-ranks} for
the full table.

\subsection{Dirichlet weight perturbation: full volatility table}

\Cref{tab:dirichlet-volatility} in \cref{sec:dirichlet} reports mean
rank, rank standard deviation, and 5th/95th percentile ranks for
all~27 countries.  See \cref{tab:dirichlet-volatility} for the full
table.

\subsection{Winsorization: full rank comparison}

\Cref{tab:winsorization} in \cref{sec:winsorization} compares
baseline and winsorized ranks for all~27 countries.  See
\cref{tab:winsorization} for the full table.
